---
title: "Bellabeat Case Study"
author: "Leopoldine Mirtil"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Disclaimer

This analysis was made from the _Bellabeat Case Study: How Can a Wellness Technology Company Play It Smart?_ offered through the Google Data Analytics Certificate program on Coursera.com. The data is from the [FitBit Fitness Tracker Data](https://www.kaggle.com/datasets/arashnic/fitbit?resource=download) on Kaggle.com and was originally uploaded by the user Möbius for public use. The data covers one month of collection from 4/12/2016 to 5/12/2016 from over 30 consenting users.


## Introduction

#### Scenario

You are a junior data analyst working on the marketing analyst team at Bellabeat, a high-tech manufacturer of health-focused products for women. Bellabeat is a successful small company, but they have the potential to become a larger player in the global smart device market. Urška Sršen, cofounder and Chief Creative Officer of Bellabeat, believes that analyzing smart device fitness data could help unlock new growth opportunities for the company. You have been asked to focus on one of Bellabeat’s products and analyze smart device data to gain insight into how consumers are using their smart devices. The insights you discover will then help guide marketing strategy for the company. You will present your analysis to the Bellabeat executive team along with your high-level recommendations for Bellabeat’s marketing strategy.


#### Stakeholders & Products

* __Stakeholders__

    + __Urška__ __Sršen:__ Bellabeat’s cofounder and Chief Creative Officer

    + __Sando__ __Mur:__ Mathematician and Bellabeat’s cofounder; key member of the Bellabeat executive team

    + __Bellabeat__ __marketing__ __analytics__ __team:__ A team of data analysts responsible for collecting, analyzing, and reporting data that helps guide Bellabeat’s marketing strategy. You joined this team six months ago and have been busy learning about Bellabeat’’s mission and business goals — as well as how you, as a junior data analyst, can help Bellabeat achieve them.


* __Products__

    + __Bellabeat__ __app:__ The Bellabeat app provides users with health data related to their activity, sleep, stress, menstrual cycle, and mindfulness habits. This data can help users better understand their current habits and make healthy decisions. The Bellabeat app connects to their line of smart wellness products.

    + __Leaf:__ Bellabeat’s classic wellness tracker can be worn as a bracelet, necklace, or clip. The Leaf tracker connects to the Bellabeat app to track activity, sleep, and stress.

    + __Time:__ This wellness watch combines the timeless look of a classic timepiece with smart technology to track user activity, sleep, and stress. The Time watch connects to the Bellabeat app to provide you with insights into your daily wellness.

    + __Spring:__ This is a water bottle that tracks daily water intake using smart technology to ensure that you are appropriately hydrated throughout the day. The Spring bottle connects to the Bellabeat app to track your hydration levels.


#### Tasks

1. What are some trends in smart device usage?
2. How could these trends apply to Bellabeat customers?
3. How could these trends help influence Bellabeat marketing strategy?


## Get Started

### Step 1 - Import Data

#### Load Packages

```{r, warning=FALSE, message=FALSE}
library(tidyr)
library(tidyverse)
library(dplyr)
library(knitr)
library(lubridate)
library(chron)
library(pastecs)
library(ggplot2)
```


#### Remove Scientific Notation Format

```{r}
options(scipen = 50) 
```


#### Set Directory & Import Data 

```{r, error=FALSE, warning=FALSE}
setwd('C:/Users/Leopoldine/Desktop/Mine/Coding Projects & Portfolio/Bellabeat Case Study/00_raw_data')

dailyActs <- read.csv('dailyActivity_merged.csv')
dailyCals <- read.csv('dailyCalories_merged.csv')
dailyInts <- read.csv('dailyIntensities_merged.csv')
dailySteps <- read.csv('dailySteps_merged.csv')
hrate_sec <- read.csv('heartrate_seconds_merged.csv')
hrCals <- read.csv('hourlyCalories_merged.csv')
hrInts <- read.csv('hourlyIntensities_merged.csv')
hrSteps <- read.csv('hourlySteps_merged.csv')
minCalsN <- read.csv('minuteCaloriesNarrow_merged.csv')  
minCalsW <- read.csv('minuteCaloriesWide_merged.csv')
minIntsN <- read.csv('minuteIntensitiesNarrow_merged.csv')
minIntsW <- read.csv('minuteIntensitiesWide_merged.csv')
minMETsN <- read.csv('minuteMETsNarrow_merged.csv')
minSleep <- read.csv('minuteSleep_merged.csv')
minStepsN <- read.csv('minuteStepsNarrow_merged.csv')
minStepsW <- read.csv('minuteStepsWide_merged.csv')
sleepDay <- read.csv('sleepDay_merged.csv')
weightLog <- read.csv('weightLogInfo_merged.csv')
```


### Step 2 - Clean Data

#### Review Daily Data Sets

##### Compare Daily Dataframes

```{r}
##dailyActs vs dailyCals
# TRUE = equal, False=not equal
all.equal(dailyActs$Calories, dailyCals$Calories) 
all.equal(dailyActs$Id, dailyCals$Id)
all.equal(dailyActs$ActivityDate, dailyCals$ActivityDay)
```


```{r}
##dailyActs vs dailySteps
all.equal(dailyActs$TotalSteps, dailySteps$StepTotal) 
all.equal(dailyActs$Id, dailySteps$Id)
all.equal(dailyActs$ActivityDate, dailySteps$ActivityDay)
```


```{r}
#dailyActs vs dailyInts
all.equal(dailyActs$Id, dailyInts$Id)
all.equal(dailyActs$ActivityDate, dailyInts$ActivityDay)
all.equal(dailyActs[7:14], dailyInts[3:10]) 
```


```{r}
##convert columns of dailyActs & dailyInts to data tables for comparison
d_Acts <- data.table::setDT(dailyActs[7:14])
d_Ints <- data.table::setDT(dailyInts[3:10])

all.equal(d_Acts, d_Ints, ignore.col.order = TRUE)# ignore column order
```

```{r, results='hide', include=FALSE}
#clean environment
#remove unneeded datasets
rm(d_Acts, d_Ints, dailyCals, dailyInts, dailySteps)
gc()
```


##### Change Data Type of Column

```{r}
#will be a common point in other data sets
dailyActs$ActivityDate <- as.Date(dailyActs$ActivityDate, '%m/%d/%Y') 
```


##### Examine Modified 'dailyActs' Dataframe

```{r}
str(dailyActs)
```


#### Review Hourly Data sets

##### Compare Hourly Dataframes

```{r}
#check Id columns
all.equal(hrCals$Id, hrInts$Id)
all.equal(hrSteps$Id, hrInts$Id)

#check Activity columns
all.equal(hrCals$ActivityHour, hrInts$ActivityHour)
all.equal(hrSteps$ActivityHour, hrInts$ActivityHour)
```


##### Merge Hourly Data Sets

```{r}
hourlyActs <- bind_cols(hrCals, hrSteps[3], hrInts[3:4])

# view data
str(hourlyActs)
```


```{r, results='hide', include=FALSE}
#clean environment and remove unneeded datasets
rm(hrCals, hrInts, hrSteps)
gc() #clear up memory
```


#### Review Minute Data Sets

##### Compare Minute Data Frames

```{r}
# check Id&ActivityMinute columns of min_narrow datasets
all.equal(minCalsN[1:2],minStepsN[1:2])
all.equal(minIntsN[1:2],minMETsN[1:2])
all.equal(minCalsN[1:2],minMETsN[1:2])

# check Id&ActivityMinute columns of min_wide datasets
all.equal(minCalsW[1:2],minStepsW[1:2])
all.equal(minIntsW[1:2],minCalsW[1:2])
```


##### Merge Minute-Narrow Dataframes

```{r}
minActsN <- bind_cols(minCalsN, minStepsN[3], minIntsN[3], minMETsN[3])
str(minActsN)
```

##### Merge Minute-Wide Dataframes

```{r}
minWide <- bind_cols(minCalsW, minIntsW[,3:62], minStepsW[,3:62]) 

# view new data set
str(minWide)
```


```{r, results='hide', include=FALSE}
#clean environment
rm(minCalsN, minStepsN, minIntsN, minMETsN, minCalsW, minIntsW, minStepsW)
gc()
```


### Step 3 - Modify Data

#### Rename Columns

```{r}
dailyActs <- rename(dailyActs, DailyActsDate=ActivityDate, DailyCalories=Calories, DailyId=Id)
sleepDay <- rename(sleepDay, SleepDateTime=SleepDay, SleepDayId=Id)
hourlyActs <- rename(hourlyActs, HourlyId=Id, HourlyCalories=Calories, HourlyStepTotal=StepTotal, HourlyTotalIntensity=TotalIntensity, HourlyAverageIntensity=AverageIntensity)
hrate_sec <- rename(hrate_sec, HRateId=Id, HRateDateTime=Time, HeartRateSec=Value)
minActsN <- rename(minActsN, MinActsId=Id, DateMinute=ActivityMinute, MinuteCalories=Calories, MinuteSteps=Steps, MinuteIntensity=Intensity, MinutesMETs=METs)
minSleep <- rename(minSleep, MinSleepId=Id, MinSleepDateTime=date, MinSleepValue=value, MinSleepLogId=logId)
weightLog <- rename(weightLog, WeightId=Id, WeightDateTime=Date, WeightLogId=LogId)
minWide <- rename(minWide, DateTime=ActivityHour)
```


#### Add New Columns

##### Add Date-Only Columns

```{r}
hourlyActs$HourlyDate <- as.Date(hourlyActs$ActivityHour, '%m/%d/%Y')
minActsN$MinuteDates <- as.Date(minActsN$DateMinute, '%m/%d/%Y')
minSleep$MinSleepDate <- as.Date(minSleep$MinSleepDateTime, '%m/%d/%Y')
sleepDay$SleepDate <- as.Date(sleepDay$SleepDateTime, '%m/%d/%Y')
weightLog$WeightDate <- as.Date(weightLog$WeightDateTime, '%m/%d/%Y')
hrate_sec$HRateDate <- as.Date(hrate_sec$HRateDateTime, '%m/%d/%Y')
minWide$Date <- as.Date(minWide$DateTime, '%m/%d/%Y %r')
```


##### Add Time-Only Columns 

```{r, warning=FALSE}
hourlyActs$Hour <- chron(times.=(format(strptime(hourlyActs$ActivityHour, '%m/%d/%Y %r'), '%H:%M:%S')))
minActsN$Minutes <- chron(times.=(format(strptime(minActsN$DateMinute, '%m/%d/%Y %r'), '%H:%M:%S')))
minSleep$SleepMinutes <- chron(times.=(format(strptime(minSleep$MinSleepDateTime, '%m/%d/%Y %r'), '%H:%M:%S')))
sleepDay$SleepTime <- chron(times.=(format(strptime(sleepDay$SleepDateTime, '%m/%d/%Y %r'), '%H:%M:%S')))
weightLog$WeightTime <- chron(times.=(format(strptime(weightLog$WeightDateTime, '%m/%d/%Y %r'), '%H:%M:%S')))
hrate_sec$HRateTime <- chron(times.=(format(strptime(hrate_sec$HRateDateTime, '%m/%d/%Y %r'), '%H:%M:%S')))
minWide$Time <- chron(times.=(format(strptime(minWide$DateTime, '%m/%d/%Y %r'), '%H:%M:%S')))
```


#### Reorganize Columns

```{r}
# move new Date & Time cols after Date/Time cols in each dataset
# make it easier to merge data sets all together
hourlyActs <- hourlyActs[, c(1, 2, 7, 8, 3:6)]
dailyActs <- dailyActs[, c(1, 2, 15, 3:14)] #move 'DailyCalories' column for easy viewing
minActsN <- minActsN[, c(1, 2, 7, 8, 3:6)]
minSleep <- minSleep[, c(1, 2, 5, 6, 3, 4)]
sleepDay <- sleepDay[, c(1, 2, 6, 7, 3:5)]
weightLog <- weightLog[, c(1, 2, 9, 10, 3:8)]
hrate_sec <- hrate_sec[, c(1, 2, 4, 5, 3)]
minWide <- minWide[, c(1, 2, 183, 184, 3:182)]
```


#### Confirm Date Ranges

```{r}
range(dailyActs$DailyActsDate)

range(hourlyActs$HourlyDate) 

range(minActsN$MinuteDates)

range(minSleep$MinSleepDate) 

range(sleepDay$SleepDate) 

range(weightLog$WeightDate) 

range(hrate_sec$HRateDate)  

range(minWide$Date) 
```


#### Remove Out of Range Dates

```{r}
# remove 4/11 data rows from minSleep
minSleep <- minSleep[minSleep[['MinSleepDate']]>='2016-04-12', ] 
#remove 5/13 data rows from minWide
minWide <- minWide[minWide[['Date']]<='2016-05-12', ]
# confirm date range
range(minSleep$MinSleepDate)
range(minWide$Date)
```


#### Count NA Values

```{r}
# count of NA values
sum(is.na(dailyActs)) 
sum(is.na(hourlyActs))
sum(is.na(minActsN)) 
sum(is.na(minSleep))
sum(is.na(sleepDay)) 
sum(is.na(weightLog)) 
sum(is.na(hrate_sec))
sum(is.na(minWide)) 
```


#### Examine Dataset for NA values

```{r}
# examine 'weightLog' data for location of all NA vals
apply(X = is.na(weightLog), MARGIN = 2, FUN = sum)
```

Decided not to remove the NA values from the data frame to avoid skewing the results.


#### Remove Column 

```{r}
minWide <- select(minWide, -c(DateTime))
```


#### Export Modified Dataframes

```{r, warning=FALSE}
setwd('C:/Users/Leopoldine/Desktop/Mine/Coding Projects & Portfolio/Bellabeat Case Study/01_tidy_data')

write.csv(dailyActs, "dailyActs.csv", row.names = FALSE)
write.csv(hourlyActs, "hourlyActs.csv", row.names = FALSE)
write.csv(hrate_sec, "hrate_sec.csv", row.names = FALSE)
write.csv(minActsN, "minActsN.csv", row.names = FALSE)
write.csv(minSleep, "minSleep.csv", row.names = FALSE)
write.csv(minWide, "minWide.csv", row.names = FALSE)
write.csv(sleepDay, "sleepDay.csv", row.names = FALSE)
write.csv(weightLog, "weightLog.csv", row.names = FALSE)
```


#### Merge Data Sets

```{r, warning=FALSE, message=FALSE}
#join dailyActs & sleepDay datsets by Id & date 
use1 <- full_join(dailyActs, sleepDay, by=join_by(DailyId==SleepDayId, DailyActsDate==SleepDate), keep=TRUE) # keep=TRUE keeps matching column
#remove unnecessary columns 
use1 <- select(use1, -c(SleepDateTime))
```


```{r, warning=FALSE, message=FALSE}
# join daily data and hourly dataset
## merge smaller dataset to larger data set to keep order in Id, Date and Time
use2 <- full_join(use1, hourlyActs, by=join_by(DailyId==HourlyId, DailyActsDate==HourlyDate, SleepTime==Hour), keep=TRUE) 
use2 <- select(use2, -c(ActivityHour))

# clear previous joined files
rm(use1, hourlyActs, dailyActs, sleepDay)
```


```{r, warning=FALSE, message=FALSE, results='hide'}
# join minute datasets together
use3 <- full_join(minActsN, minSleep, by=join_by(MinActsId==MinSleepId, MinuteDates==MinSleepDate, Minutes==SleepMinutes), keep=TRUE)
use3 <- select(use3, -c(DateMinute, MinSleepDateTime))
rm(minActsN, minSleep)
gc()

# combine use2 &  use3
use4 <- full_join(use2, use3, by=join_by(DailyId==MinActsId, DailyActsDate==MinuteDates, Hour==Minutes), keep=TRUE)
rm(use2, use3)
gc()

### will add remaining data sets separately to primary table ##################
# join use4 + hrate
use5 <- full_join(use4, hrate_sec, by=join_by(DailyId==HRateId, DailyActsDate==HRateDate, Minutes==HRateTime), keep=TRUE) 
use5 <- select(use5, -c(HRateDateTime))

# join use4.5 + weight Log
smart_dev_use <- full_join(use5, weightLog, by=join_by(DailyId==WeightId, DailyActsDate==WeightDate, HRateTime==WeightTime), keep=TRUE) 
smart_dev_use <- select(smart_dev_use, -c(WeightDateTime))
rm(use4, use5, hrate_sec, weightLog) #clear environment
gc()
```


#### Export Merged Data Frame

```{r, warning=FALSE, message=FALSE, results='hide'}
# set directory
setwd('C:/Users/Leopoldine/Desktop/Mine/Coding Projects & Portfolio/Bellabeat Case Study/01_tidy_data')

#export final merged file as precaution
write.csv(smart_dev_use, 'smart_dev_use.csv', row.names = FALSE)
```


```{r, echo=FALSE, results='hide', include=FALSE}
# clear up memory
gc()
```


### Step 4- Further Modify Combined Dataset

#### Inspect Merged Data Frame

```{r}
str(smart_dev_use)
```


#### Create Primary Id column

```{r}
# want one singular Id col in dataset
# find way to add missing 'Ids' to 'DailyId' col if present in others
smartDev <- smart_dev_use  

# Primary Id column == DailyId
## Id columns: SleepDayId, HourlyId, MinActsId, HRateId, WeightId, MinSleepId 
# will use all except SleepDayId (in range)
smartDev$DailyId <- ifelse(is.na(smartDev$DailyId), smartDev$MinActsId, smartDev$DailyId)
smartDev$DailyId <- ifelse(is.na(smartDev$DailyId), smartDev$HourlyId, smartDev$DailyId)
smartDev$DailyId <- ifelse(is.na(smartDev$DailyId), smartDev$HRateId, smartDev$DailyId)
smartDev$DailyId <- ifelse(is.na(smartDev$DailyId), smartDev$MinSleepId, smartDev$DailyId)
smartDev$DailyId <- ifelse(is.na(smartDev$DailyId), smartDev$WeightId, smartDev$DailyId)

## confirm no more NA values in Id
sum(is.na(smartDev$DailyId))
```


```{r, echo=FALSE, results='hide', include=FALSE}
# clear up memory
gc()
```


#### Create Primary Date Column 

```{r}
# Primary Column: DailyActsDate
# Date columns: SleepDate, HourlyDate, MinuteDates, MinSleepDate, HRateDate, WeightDate
# use all except SleepDate(already in range)
smartDev <- smartDev %>% mutate(DailyActsDate=coalesce(DailyActsDate, MinuteDates))
smartDev <- smartDev %>% mutate(DailyActsDate=coalesce(DailyActsDate, HourlyDate))
smartDev <- smartDev %>% mutate(DailyActsDate=coalesce(DailyActsDate, HRateDate))
smartDev <- smartDev %>% mutate(DailyActsDate=coalesce(DailyActsDate, MinSleepDate))
smartDev <- smartDev %>% mutate(DailyActsDate=coalesce(DailyActsDate, WeightDate))

# remove date columns
smartDev <- select(smartDev, -c(SleepDate, HourlyDate, MinuteDates, MinSleepDate, HRateDate, WeightDate))

## confirm no more NA values in Date
sum(is.na(smartDev$DailyActsDate))
```

#### Create Primary Time Column

```{r}
smartDev$Time <- smartDev$SleepTime

# merge other time columns # times: Hour, Minutes, SleepMinutes, HRateTime, WeightTime (covers largest time incl SleepTime)
# merge time columns
smartDev <- smartDev %>% mutate(Time=coalesce(Time, Minutes))
smartDev <- smartDev %>% mutate(Time=coalesce(Time, Hour))
smartDev <- smartDev %>% mutate(Time=coalesce(Time, HRateTime))
smartDev <- smartDev %>% mutate(Time=coalesce(Time, SleepMinutes))
smartDev <- smartDev %>% mutate(Time=coalesce(Time, WeightTime))

# remove columns
smartDev <- select(smartDev, -c(SleepTime, Hour, Minutes, SleepMinutes, HRateTime, WeightTime))

# move new 'Time' column to after date
smartDev <- smartDev[, c(1, 2, 42, 3:41)]
```


#### Rename Modified Column

```{r, results='hide'}
# remove other Id columns
smartDev <- select(smartDev, -c(SleepDayId, HourlyId, MinActsId, HRateId, WeightId, MinSleepId, MinSleepLogId, WeightLogId))

# Rename columns
smartDev <- rename(smartDev, Id=DailyId, Date=DailyActsDate)

# clear environment
rm(smart_dev_use)
gc()
```


#### View Merged Dataframe

```{r}
str(smartDev)
```


#### Export Final Merged Dataframe

```{r, warning=FALSE, message=FALSE}
setwd('C:/Users/Leopoldine/Desktop/Mine/Coding Projects & Portfolio/Bellabeat Case Study/01_tidy_data')

#export final merged file
write.csv(smartDev, 'smartDev.csv', row.names = FALSE)
```


### Step 5 - Transpose Dataframes

#### Pivot Smart Features Dataframe

```{r, warning=FALSE}
# create new dataframe in case of error
smart_d2 <- smartDev

#transpose data while keeping 3 cols: Id(inc), Dates, & Time
mod_smart <- 
smart_d2 %>%
  pivot_longer(cols=c(DailyCalories:MinSleepValue, HeartRateSec:BMI), 
               names_to = 'SmartFeatures', 
               values_to = 'Values', 
               values_drop_na = TRUE)

#view pivoted data
head(mod_smart) 
```


```{r, results='hide', echo=FALSE, include=FALSE}
gc() # clear up storage/memory
```


##### Pivot Minute-Wide Features Dataframe

```{r, warning=FALSE}
#pivot data, keeping Id, Date, and Time cols
## split Cals00-59, Steps00-59, Ints00-59 
mod_minWide <-
minWide %>%
pivot_longer(
    cols=!c(Id, Date, Time),
    names_to = c('Features', '.value'), 
    names_pattern = '(.*)(.\\d)')

#view pivoted data
head(mod_minWide)
```


```{r, warning=FALSE}
### further transpose to create new Minutes column 
mod_mins <-
mod_minWide %>%
pivot_longer(cols = !c(Id, Date, Time, Features),
             names_to = 'Minutes',
             values_to = 'Value') %>%
  rename(Hour=Time) 

#reorder  columns to move 'Minutes' after Hour
mod_mins <- mod_mins[,c(1:3,5,4,6)] 
mod_mins$Minutes <- chron(times.=(format(strptime(mod_mins$Minutes, '%M'), '%H:%M:%S')))

#view pivoted data
head(mod_mins)
```


```{r, warning=FALSE}
#merge Minutes and Hours into new 'Time' column
mod_mins <-
mod_mins %>% 
  mutate(Hour=hours(Hour), Minutes=minutes(Minutes)) %>%
  unite(Time, Hour, Minutes, sep=':') %>%
  mutate(Time=chron(times.=(format(strptime(Time, '%H:%M'), '%H:%M:%S'))))

# view final modified data frame
head(mod_mins)
```


##### Check for NA Values 

```{r}
#expect a fair amount as users did not use all features at all times
sum(is.na(mod_smart)) 

# should be zero, same as before
sum(is.na(mod_mins))
```


```{r, results='hide', include=FALSE}
#clean environment and memory
rm(minWide, smart_d2, mod_minWide, smartDev)
gc() 
```


#### Export Modified File

```{r, warning=FALSE}
setwd('C:/Users/Leopoldine/Desktop/Mine/Coding Projects & Portfolio/Bellabeat Case Study/01_tidy_data')

write.csv(mod_smart, 'mod_smart.csv', row.names = FALSE)
write.csv(mod_mins, 'mod_mins.csv', row.names = FALSE)
```


### Step 6 - Analyze Data

#### Descriptive Analysis of Smart Features

```{r}
round(stat.desc(mod_smart), 2)
```

#### Summary Statistics of Smart Features

```{r}
summary(mod_smart)
```

#### Descriptive Analysis of Common Features

```{r}
round(stat.desc(mod_mins), 2)
```

#### Summary Statistics of Common Features

```{r}
summary(mod_mins)
```


#### User Count

##### Distinct Count of Users per Features

```{r}
mod_smart %>%
  group_by(SmartFeatures)%>%
  summarise(UserCount=length(unique(Id))) %>%
  arrange(desc(UserCount))
```


##### Distinct Count of Users per Common Features  

```{r}
mod_mins %>%
  group_by(Features)%>%
  summarise(UserCount=length(unique(Id))) %>%
  arrange(desc(UserCount))
```


##### Total Count of Manual Report Usage

```{r}
mod_smart %>% count(IsManualReport, sort = TRUE) %>% rename(Total=n) %>% drop_na()
```


##### Number of Manual Report Users

```{r}
mod_smart %>%
  group_by(IsManualReport)%>%
  summarise(WeightUserCount=length(unique(Id))) %>%
  arrange(desc(WeightUserCount)) %>%
  drop_na() 
```


#### Frequency Count

##### Total Count of Smart Feature Usage

```{r, warning=FALSE, message=FALSE}
mod_smart %>% count(SmartFeatures, sort = TRUE) %>% rename(Frequency=n)
```


#### Common Features Frequency 

```{r}
mod_mins %>% count(Features, sort = TRUE) %>% rename(Frequency=n)
```


#### Date

##### Total Daily Count of Smart Features 

```{r, message=FALSE}
mod_smart %>% 
  group_by(Date) %>%
  summarise(Count = n()) %>%
  arrange(Date, desc(Count))
```


##### Total Daily Count of Smart Features Frequency 

```{r, warning=FALSE, message=FALSE}
mod_smart %>% 
  group_by(Date, SmartFeatures) %>%
  summarise(Total_Count = n()) %>%
  arrange(Date, desc(Total_Count))
```


##### Total Daily Count of Common Features 

```{r, message=FALSE}
mod_mins %>% 
  group_by(Date) %>%
  summarise(Count = n()) %>%
  arrange(Date)
```


##### Total Daily Count of Common Features Frequency 

```{r, warning=FALSE, message=FALSE}
mod_mins %>% 
  group_by(Date, Features) %>%
  summarise(Total_Count = n()) %>%
  arrange(Date)
```


##### Total Weekday Count of Smart Features

```{r}
mod_smart %>% 
  mutate(Weekday=wday(Date, label=TRUE)) %>%
  group_by(Weekday) %>%
  summarise(Frequency = n()) %>%
  arrange(Weekday, desc(Frequency)) 
```



##### Weekday Frequency of Smart Features 

```{r, message=FALSE}
mod_smart %>% 
  mutate(Weekday=wday(Date, label=TRUE)) %>%
  group_by(Weekday, SmartFeatures) %>%
  summarise(Frequency = n()) %>%
  arrange(Weekday, desc(Frequency)) 
```



##### Daily Average of Common Features 

```{r, warning=FALSE, message=FALSE}
mod_mins %>%
  group_by(Date, Features) %>%
  summarise(Average = mean(Value)) %>%
  arrange(Date, desc(Average))
```


##### Weekday Frequency of Common Features

```{r, warning=FALSE, message=FALSE}
mod_mins %>% 
  mutate(Weekday=wday(Date, label=TRUE)) %>%
  group_by(Weekday) %>%
  summarise(Frequency = n()) %>%
  arrange(Weekday) 
```


##### Weekday Average of Common Features

```{r, warning=FALSE, message=FALSE}
mod_mins %>% 
  mutate(Weekday=wday(Date, label=TRUE)) %>%
  group_by(Weekday, Features) %>%
  summarise(Mean = round(mean(Value), 2)) %>%
  arrange(Weekday, desc(Mean)) 
```


#### Time 

##### Hourly Frequency of Smart Features

```{r, warning=FALSE, message=FALSE}
## total feature count per hour
mod_smart %>%
  mutate(Hour=hours(Time)) %>%
  group_by(Hour, SmartFeatures) %>%
  summarise(Total_Count=n()) %>%
  arrange(Hour, desc(Total_Count))%>%
  drop_na() 
```


##### Hourly Frequency of Common Features

```{r, warning=FALSE, message=FALSE}
mod_mins %>%
  mutate(Hour=hours(Time)) %>%
  group_by(Hour, Features) %>%
  summarise(Total_Count = n()) %>%
  arrange(Hour, desc(Total_Count))
```


##### Hourly Average of Common Features

```{r, warning=FALSE, message=FALSE}
mod_mins %>%
  mutate(Hour=hours(Time)) %>%
  group_by(Hour, Features) %>%
  summarise(Average = round(mean(Value), 2)) %>%
  arrange(Hour, desc(Average))
```


##### Minute Frequency of Smart Features 

```{r, warning=FALSE, message=FALSE}
mod_smart %>%
  mutate(Minutes=minutes(Time)) %>%
  group_by(Minutes, SmartFeatures) %>%
  summarise(Total_Count=n()) %>%
  drop_na() %>%
  arrange(Minutes, desc(Total_Count))
```

##### Minute Frequency of Common Features 

```{r, warning=FALSE, message=FALSE}
mod_mins %>%
  mutate(Minutes=minutes(Time)) %>%
  group_by(Minutes, Features) %>%
  summarise(Total_Count = n()) %>%
  arrange(Minutes, desc(Total_Count))
```


##### Minutes Average of Common Features 

```{r, warning=FALSE, message=FALSE}
mod_mins %>%
  mutate(Minutes=minutes(Time)) %>%
  group_by(Minutes, Features) %>%
  summarise(Average = round(mean(Value), 2)) %>%
  arrange(Minutes, desc(Average)) 
```


#### Rank Analysis

##### Top 5 Most Popular Features

```{r, warning=FALSE, message=FALSE}
mod_smart %>%
  group_by(SmartFeatures) %>%
  summarise(Total=n()) %>%
  arrange(desc(Total)) %>%
  slice(1:5)
```

##### Top 5 Least Popular Features

```{r, warning=FALSE, message=FALSE}
mod_smart %>%
  group_by(SmartFeatures) %>%
  summarise(Total=n()) %>%
  arrange(Total) %>%
  slice(1:5)
```

#### Step 7 - Visualizations

```{r, warning=FALSE, message=FALSE, echo=FALSE}
mod_smart %>%
  group_by(SmartFeatures)%>%
  summarise(User_Count=length(unique(Id))) %>%
  arrange(desc(User_Count)) %>%
  ggplot(mapping=aes(x=fct_reorder(SmartFeatures,-User_Count), y=User_Count)) +
  geom_col(position = "dodge", fill='steelblue') + 
   ggtitle('User Count per Features') +
  xlab('Smart Features') +
  ylab('Number of Users') +
  coord_flip() +
  theme(legend.position = "none") 
```

Most of the smart features were utilized by all 33 users. There were 9 smart features that had under 25 users. 


```{r, warning=FALSE, message=FALSE, echo=FALSE}
mod_smart %>%
  group_by(SmartFeatures) %>%
  summarise(Total_Count=n()) %>%
  arrange(desc(Total_Count)) %>%
  ggplot(mapping=aes(x=fct_reorder(SmartFeatures, +Total_Count), y=Total_Count)) +
  geom_col(position = "dodge", fill='lightseagreen') + 
  xlab('Smart Features') +
  ylab('Total Count') +
  ggtitle('Frequency of Smart Features') +
  scale_y_continuous(labels = scales::comma) + 
  coord_flip() +
  theme(legend.position = "none") 
```

There was a large disparity in the frequency of use among the smart features. Five of the features recorded a over 1,250,000 total uses at the end of the month. The remaining features had under 250,000 frequency total.




```{r, warning=FALSE, message=FALSE, echo=FALSE}
mod_smart %>%
  group_by(SmartFeatures) %>%
  summarise(Total_Count=n()) %>%
  arrange(desc(Total_Count)) %>%
  slice(1:5) %>%
  ggplot(mapping=aes(x=fct_reorder(SmartFeatures, +Total_Count), y=Total_Count)) +
  geom_col(position = "dodge", fill='maroon') + 
  xlab('Smart Features') +
  ylab('Total Count') +
  ggtitle('Top 5 Most Popular Features') +
  geom_text(aes(label=format(Total_Count, big.mark = ","), y=Total_Count, 
            hjust=ifelse(Total_Count < mean(range(Total_Count)), -0.05, 1.1))) +
  scale_y_continuous(labels = scales::comma) + 
  coord_flip() +
  theme(legend.position = "none") 
```

The top 5 most frequently used smart features among the users possess a time measure (seconds, minutes and hourly). As the time factor of the features increased (from seconds to hourly), the frequency decreased. 



```{r, warning=FALSE, message=FALSE, echo=FALSE}
mod_smart %>%
  group_by(SmartFeatures) %>%
  summarise(Total_Count=n()) %>%
  arrange(Total_Count) %>%
  slice(1:5) %>%
  ggplot(mapping=aes(x=fct_reorder(SmartFeatures,+Total_Count), y=Total_Count)) +
  geom_col(stat='identity', position = "dodge", fill='gold3') +
  geom_text(aes(label=Total_Count, y=Total_Count, 
            hjust=ifelse(Total_Count < mean(range(Total_Count)), -0.15, 1.1))) +
  xlab('Smart Features') +
  ylab('Total Count') +
  ggtitle('Top 5 Least Popular Features') +
  coord_flip() +
 theme(legend.position = "none") 
```

The least used features have a total use count under 500 among the users. These features are compromised of the weight features and had less than 70 use count, with the lowest being Fat with only 2 over the course of the month.


#### Weekday Frequency 

```{r, message=FALSE, echo=FALSE}
mod_smart %>% 
  mutate(Weekday=wday(Date, label=TRUE)) %>%
  group_by(Weekday) %>%
  summarise(Frequency = n()) %>%
  arrange(Weekday) %>%
    ggplot(aes(x=Weekday, y=Frequency)) + 
    geom_col(fill='cadetblue4') +
    ggtitle('Weekday Frequency of Smart Features') +
    scale_y_continuous(labels = scales::comma)
```

The frequency of device usage was consistently over 1,000,000 throughout the weekdays. The highest usage was seen midweek, from Tuesday to Thursday. 


```{r, message=FALSE, echo=FALSE}
mod_mins %>% 
  mutate(Weekday=wday(Date, label=TRUE)) %>%
  group_by(Weekday, Features) %>%
  summarise(Frequency = n()) %>%
  arrange(Weekday) %>%
    ggplot(aes(x=Weekday, y=Frequency, fill=Features)) + 
    geom_col() +
    ggtitle('Weekday Frequency of Common Features') +
    scale_y_continuous(labels = scales::comma) +
    theme(legend.position = "none") +
    facet_grid(vars(Features))
```

These three features, used by all sampled users, were consistently above 150,000 use count throughout the week. The highest point of use across all three features was seen on Wednesdays and Thursdays, with  a frequency over 200,000.


#### Time Frequency of Smart Features

```{r, warning=FALSE, message=FALSE, echo=FALSE}
library(patchwork) # side by side plot

# count by hour
m1 <- mod_smart %>% 
  mutate(Hour=hours(Time)) %>%
  group_by(Hour) %>%
  summarise(Frequency = n()) %>%
  arrange(Hour) %>%
    ggplot(aes(x=Hour, y=Frequency)) +
    geom_area(fill='royalblue3', alpha=0.5) +
    geom_line(color='royalblue3', size=1) +
    ggtitle('Hourly Frequency of Smart \n Features') +
    scale_y_continuous(labels = scales::comma)

#count by minutes
m2 <- mod_smart %>% 
  mutate(Minutes=minutes(Time)) %>%
  group_by(Minutes) %>%
  summarise(Frequency = n()) %>%
  arrange(Minutes) %>%
    ggplot(aes(x=Minutes, y=Frequency)) +
    geom_area(fill='firebrick', alpha=0.5) +
    geom_line(color='firebrick', size=1) +
    ggtitle('Minute Frequency of Smart \n Features') +
    scale_y_continuous(labels = scales::comma)

m1+m2
```

The hourly frequency of the smart features was greater than that of the minutes, but not as consistent. The smart features had a frequency of use over 300,000 across the hours. The highest points of use occurred between hours 9 and  19 with over 325,000 uses. The minute frequency had the highest point of use at minute '0' at a little over 225,000. The remaining minutes had a consistent frequency over 125,000 uses.


#### Time Frequency of Common Features

```{r, message=FALSE, echo=FALSE}
library(patchwork)

h1 <- mod_mins %>%
  mutate(Hours=hours(Time)) %>%
  group_by(Hours, Features) %>%
  summarise(Total_Count= n()) %>%
  ggplot(aes(x=Hours, y=Total_Count, group=Features, color=Features)) +
    geom_smooth(method = "loess",
              se = FALSE,
              formula = 'y ~ x',
              span = 0.8) +
    stat_smooth(se=FALSE, geom="area",
              method = 'loess', alpha=0.5,
              span = 0.8,aes(fill=Features)) +
    ggtitle('Hourly Frequency of Common \n Features') +
    theme(legend.position = "none") +
    scale_y_continuous(labels = scales::comma) +
    ylab("Frequency") +
    facet_grid(vars(Features)) 


h2 <- mod_mins %>%
  mutate(Minutes=minutes(Time)) %>%
  group_by(Minutes, Features) %>%
  summarise(Total_Count= n()) %>%
  ggplot(aes(x=Minutes, y=Total_Count, group=Features, color=Features)) +
    geom_smooth(method = "loess",
              se = FALSE,
              formula = 'y ~ x',
              span = 0.8) +
    stat_smooth(se=FALSE, geom="area",
              method = 'loess', alpha=0.5,
              span = 0.8,aes(fill=Features)) +
    ggtitle('Minute Frequency of Common \n Features') +
    theme(legend.position = "none") +
    scale_y_continuous(labels = scales::comma) +
    ylab("Frequency") +
    facet_grid(vars(Features)) 

h1 + h2
```

The frequency of the three main features are consistent across the hours and minutes. The hourly frequency was identical to each each feature and consistently over 50,000 uses from hour '0' to '23'. The minute frequency was consistently over 20,000 across the minutes.



### Top Recommendations

**Drop Least Popular Features**

Bellabeat should consider dropping the five least popular smart features. This would allow the company to shift their resources onto their more profitable features and better market their products to future customers. The company should also consider either discontinuing the manual reports for weight smart features or modifying it in order to apply to other popular features.


**Develop 'Heart' Related Features**

Since the 'HeartRateSec' feature saw the most frequent use by all sampled users the company should develop more heart related features such as blood pressure or a pulse monitor. They can also consider developing a device specifically related to cardio exercises to attract potential customers. 


**Develop 'METS' Related Features**

The company should develop more 'METs' related features as it was among the top five most used features over the month and all 33 users made use of it. This would also expand the available minute features (Calories, Intensity, and Steps) used by all users from 3  to 4. The increased marketing as a result of the developments would attract more customers.
