---
title: "Bellabeat Case Study"
author: "Leopoldine Mirtil"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Disclaimer

This analysis was made from the _Bellabeat Case Study: How Can a Wellness Technology Company Play It Smart?_, offered through the Google Data Analytics Certificate program on Coursera.com. The data is from the [FitBit Fitness Tracker Data](https://www.kaggle.com/datasets/arashnic/fitbit?resource=download) from Kaggle.com and was originally uploaded by the user Möbius for public use. The data covers one month of collection from 4/12/2016 to 5/12/2016 from over 30 consenting users.


## Introduction

#### Scenario

You are a junior data analyst working on the marketing analyst team at Bellabeat, a high-tech manufacturer of health-focused products for women. Bellabeat is a successful small company, but they have the potential to become a larger player in the global smart device market. Urška Sršen, cofounder and Chief Creative Officer of Bellabeat, believes that analyzing smart device fitness data could help unlock new growth opportunities for the company. You have been asked to focus on one of Bellabeat’s products and analyze smart device data to gain insight into how consumers are using their smart devices. The insights you discover will then help guide marketing strategy for the company. You will present your analysis to the Bellabeat executive team along with your high-level recommendations for Bellabeat’s marketing strategy.


#### Tasks

1. What are some trends in smart device usage?
2. How could these trends apply to Bellabeat customers?
3. How could these trends help influence Bellabeat marketing strategy?


## Get to Work

### Step 1 - Import Data

#### Load Packages

```{r, warning=FALSE, message=FALSE}
library(tidyr)
library(tidyverse)
library(dplyr)
library(knitr)
library(lubridate)
library(chron)
library(pastecs)
library(ggplot2)
```

I chose these specific packages to enable: data manipulation, documentation, descriptive statistics and visualization.


```{r, results=FALSE, echo=FALSE}
#remove scientific notation format
options(scipen = 50) 
```


#### Set Directory & Import Data 

```{r, error=FALSE, warning=FALSE}
setwd('C:/Users/Leopoldine/Desktop/Mine/Coding Projects & Portfolio/Bellabeat Case Study/00_raw_data')

dailyActs <- read.csv('dailyActivity_merged.csv')
dailyCals <- read.csv('dailyCalories_merged.csv')
dailyInts <- read.csv('dailyIntensities_merged.csv')
dailySteps <- read.csv('dailySteps_merged.csv')
hrate_sec <- read.csv('heartrate_seconds_merged.csv')
hrCals <- read.csv('hourlyCalories_merged.csv')
hrInts <- read.csv('hourlyIntensities_merged.csv')
hrSteps <- read.csv('hourlySteps_merged.csv')
minCalsN <- read.csv('minuteCaloriesNarrow_merged.csv')  
minCalsW <- read.csv('minuteCaloriesWide_merged.csv')
minIntsN <- read.csv('minuteIntensitiesNarrow_merged.csv')
minIntsW <- read.csv('minuteIntensitiesWide_merged.csv')
minMETsN <- read.csv('minuteMETsNarrow_merged.csv')
minSleep <- read.csv('minuteSleep_merged.csv')
minStepsN <- read.csv('minuteStepsNarrow_merged.csv')
minStepsW <- read.csv('minuteStepsWide_merged.csv')
sleepDay <- read.csv('sleepDay_merged.csv')
weightLog <- read.csv('weightLogInfo_merged.csv')
```

I set the directory first to make it easier to import all the data files without having to include the full file path every time.


### Step 2 - Clean Data

#### Review Daily Data Sets

I noticed that the 'dailyActs' file had similar columns to the other daily data sets. I decided to first confirm if they were identical, and if so, remove those files. 


##### Compare Daily Dataframes

```{r}
##dailyActs vs dailyCals
# TRUE = equal, False=not equal
all.equal(dailyActs$Calories, dailyCals$Calories) 
all.equal(dailyActs$Id, dailyCals$Id)
all.equal(dailyActs$ActivityDate, dailyCals$ActivityDay)
```


```{r}
##dailyActs vs dailySteps
all.equal(dailyActs$TotalSteps, dailySteps$StepTotal) 
all.equal(dailyActs$Id, dailySteps$Id)
all.equal(dailyActs$ActivityDate, dailySteps$ActivityDay)
```


```{r}
#dailyActs vs dailyInts
all.equal(dailyActs$Id, dailyInts$Id)
all.equal(dailyActs$ActivityDate, dailyInts$ActivityDay)
```


```{r}
#dailyActs vs dailyInts
##convert columns into data tables for comparison
d_Acts <- data.table::setDT(dailyActs[7:14])
d_Ints <- data.table::setDT(dailyInts[3:10])

all.equal(d_Acts, d_Ints, ignore.col.order = TRUE)
```

I had to convert the remaining columns in 'dailyActs' & 'dailyInts' into data tables, due to different column orders, in order to perform a comparison.

```{r}
#compare similar columns 'TotalDistance' vs 'TrackerDistance'
tod <- data.table::setDT(dailyActs[5])
trd <- data.table::setDT(dailyActs[6])

all.equal(tod, trd)

#remove a column
dailyActs <- select(dailyActs, -c(TotalDistance))
```

I removed the 'TotalDistance' column after confirming it was identical to the 'TrackerDistance' column, other than the column name. There was no point in having a duplicate column.


```{r, results='hide', include=FALSE}
#clean environment
#remove unneeded datasets
rm(d_Acts, d_Ints, dailyCals, dailyInts, dailySteps, tod, trd)
gc()
```


##### Change Data Type of Column

```{r}
dailyActs$ActivityDate <- as.Date(dailyActs$ActivityDate, '%m/%d/%Y') 
```

After removing the unnecessary 'daily' files, I changed the data type and format of the 'ActivityDate' column as I will use it as a key point when merging the data sets later on.


#### Review Hourly Data sets

##### Compare Hourly Dataframes

```{r}
#check Id columns
all.equal(hrCals$Id, hrInts$Id)
all.equal(hrSteps$Id, hrInts$Id)

#check Activity columns
all.equal(hrCals$ActivityHour, hrInts$ActivityHour)
all.equal(hrSteps$ActivityHour, hrInts$ActivityHour)
```

I confirmed that the 'Id' and 'ActivityHour' columns were identical in order to use both as keys when merging the data files.


##### Merge Hourly Data Sets

```{r}
hourlyActs <- bind_cols(hrCals, hrSteps[3], hrInts[3:4])

# view data
str(hourlyActs)
```


```{r, results='hide', include=FALSE}
#clean environment and remove unneeded datasets
rm(hrCals, hrInts, hrSteps)
gc() #clear up memory
```


#### Review Minute Data Sets

##### Compare Minute Data Frames

```{r}
# check Id&ActivityMinute columns 
all.equal(minCalsN[1:2],minStepsN[1:2])
all.equal(minIntsN[1:2],minMETsN[1:2])
all.equal(minCalsN[1:2],minMETsN[1:2])

all.equal(minCalsW[1:2],minStepsW[1:2])
all.equal(minIntsW[1:2],minCalsW[1:2])
```

Same as above, I confirmed that the 'Id' and 'ActivityMinute' columns were identical before merging the files.


##### Merge Minute-Narrow Dataframes

```{r}
minActsN <- bind_cols(minCalsN, minStepsN[3], minIntsN[3], minMETsN[3])
str(minActsN)
```

##### Merge Minute-Wide Dataframes

```{r}
minWide <- bind_cols(minCalsW, minIntsW[,3:62], minStepsW[,3:62]) 
```


```{r, results='hide', include=FALSE}
#clean environment
rm(minCalsN, minStepsN, minIntsN, minMETsN, minCalsW, minIntsW, minStepsW)
gc()
```


### Step 3 - Modify Data

#### Rename Columns

```{r}
dailyActs <- rename(dailyActs, DailyActsDate=ActivityDate, DailyCalories=Calories, DailyId=Id)
sleepDay <- rename(sleepDay, SleepDateTime=SleepDay, SleepDayId=Id)
hourlyActs <- rename(hourlyActs, HourlyId=Id, HourlyCalories=Calories, HourlyStepTotal=StepTotal, HourlyTotalIntensity=TotalIntensity, HourlyAverageIntensity=AverageIntensity)
hrate_sec <- rename(hrate_sec, HRateId=Id, HRateDateTime=Time, HeartRateSec=Value)
minActsN <- rename(minActsN, MinActsId=Id, DateMinute=ActivityMinute, MinuteCalories=Calories, MinuteSteps=Steps, MinuteIntensity=Intensity, MinutesMETs=METs)
minSleep <- rename(minSleep, MinSleepId=Id, MinSleepDateTime=date, MinSleepValue=value, MinSleepLogId=logId)
weightLog <- rename(weightLog, WeightId=Id, WeightDateTime=Date, WeightLogId=LogId)
minWide <- rename(minWide, DateTime=ActivityHour)
```

I changed the names of these columns to be more descriptive and for easier identification when combining all the files.


#### Add New Columns

##### Add Date-Only Columns

```{r}
hourlyActs$HourlyDate <- as.Date(hourlyActs$ActivityHour, '%m/%d/%Y')
minActsN$MinuteDates <- as.Date(minActsN$DateMinute, '%m/%d/%Y')
minSleep$MinSleepDate <- as.Date(minSleep$MinSleepDateTime, '%m/%d/%Y')
sleepDay$SleepDate <- as.Date(sleepDay$SleepDateTime, '%m/%d/%Y')
weightLog$WeightDate <- as.Date(weightLog$WeightDateTime, '%m/%d/%Y')
hrate_sec$HRateDate <- as.Date(hrate_sec$HRateDateTime, '%m/%d/%Y')
minWide$Date <- as.Date(minWide$DateTime, '%m/%d/%Y %r')
```


##### Add Time-Only Columns 

```{r, warning=FALSE}
hourlyActs$Hour <- chron(times.=(format(strptime(hourlyActs$ActivityHour, '%m/%d/%Y %r'), '%H:%M:%S')))
minActsN$Minutes <- chron(times.=(format(strptime(minActsN$DateMinute, '%m/%d/%Y %r'), '%H:%M:%S')))
minSleep$SleepMinutes <- chron(times.=(format(strptime(minSleep$MinSleepDateTime, '%m/%d/%Y %r'), '%H:%M:%S')))
sleepDay$SleepTime <- chron(times.=(format(strptime(sleepDay$SleepDateTime, '%m/%d/%Y %r'), '%H:%M:%S')))
weightLog$WeightTime <- chron(times.=(format(strptime(weightLog$WeightDateTime, '%m/%d/%Y %r'), '%H:%M:%S')))
hrate_sec$HRateTime <- chron(times.=(format(strptime(hrate_sec$HRateDateTime, '%m/%d/%Y %r'), '%H:%M:%S')))
minWide$Time <- chron(times.=(format(strptime(minWide$DateTime, '%m/%d/%Y %r'), '%H:%M:%S')))
dailyActs$DailyTime <- chron(times.=(as.numeric(NA))) 
```

I added date and time columns to the data sets to use as primary keys when merging all the files together.


#### Reorganize Columns

```{r}
# move new Date & Time cols 
hourlyActs <- hourlyActs[, c(1, 2, 7, 8, 3:6)]
minActsN <- minActsN[, c(1, 2, 7, 8, 3:6)]
minSleep <- minSleep[, c(1, 2, 5, 6, 3, 4)]
sleepDay <- sleepDay[, c(1, 2, 6, 7, 3:5)]
weightLog <- weightLog[, c(1, 2, 9, 10, 3:8)]
hrate_sec <- hrate_sec[, c(1, 2, 4, 5, 3)]
minWide <- minWide[, c(1, 2, 183, 184, 3:182)]

#move 'DailyCalories' and Time
dailyActs <- dailyActs[, c(1, 2, 15, 14, 3:13)] 
```

I moved the new date and time columns for readability and to make it easier to merge the data sets all together. I also took the time to move the 'DailyCalories' column from the end of the 'dailyActs' data set.


#### Confirm Date Ranges

The data was collected over a one month period from 4/12/2016 to 5/12/2016. I decided to verify that all dates of all the data sets were in range.

```{r}
range(dailyActs$DailyActsDate)

range(hourlyActs$HourlyDate) 

range(minActsN$MinuteDates)

range(minSleep$MinSleepDate) 

range(sleepDay$SleepDate) 

range(weightLog$WeightDate) 

range(hrate_sec$HRateDate)  

range(minWide$Date) 
```

The only data sets containing out of range dates are 'minSleep' (4/11/2016 - 5/12/2016) and 'minWide' (4/13/2016- 5/13/2016). 


#### Remove Out of Range Dates

```{r}
# remove 4/11 data rows from minSleep
minSleep <- minSleep[minSleep[['MinSleepDate']]>='2016-04-12', ] 
#remove 5/13 data rows from minWide
minWide <- minWide[minWide[['Date']]<='2016-05-12', ]
# confirm date range
range(minSleep$MinSleepDate)
range(minWide$Date)
```

I decided to remove the out of range dates from both data sets in order to keep the dates within range.


```{r, results='hide', include=FALSE}
#calculate missing values
sum(is.na(dailyActs)) 
sum(is.na(hourlyActs))
sum(is.na(minActsN)) 
sum(is.na(minSleep))
sum(is.na(sleepDay)) 
sum(is.na(weightLog)) 
sum(is.na(hrate_sec))
sum(is.na(minWide)) 
```


```{r, results='hide', include=FALSE}
# examine 'weightLog' data for location of all NA vals
apply(X = is.na(weightLog), MARGIN = 2, FUN = sum)
#won't remove NA values to avoid possibility of skewing results
```


#### Convert Logical Values to Integers 

```{r}
weightLog$IsManualReport <- as.integer(as.logical(weightLog$IsManualReport))
```

I converted the logical values of the 'IsManualReport' column into integers where TRUE=1 and FALSE=0. I did this to enable easier analysis of this feature.

#### Remove Column 

```{r}
minWide <- select(minWide, -c(DateTime))
```

I removed this column because it was unneeded after adding the date and time columns.

#### Export Modified Dataframes

```{r, warning=FALSE}
setwd('C:/Users/Leopoldine/Desktop/Mine/Coding Projects & Portfolio/Bellabeat Case Study/01_tidy_data')

write.csv(dailyActs, "dailyActs.csv", row.names = FALSE)
write.csv(hourlyActs, "hourlyActs.csv", row.names = FALSE)
write.csv(hrate_sec, "hrate_sec.csv", row.names = FALSE)
write.csv(minActsN, "minActsN.csv", row.names = FALSE)
write.csv(minSleep, "minSleep.csv", row.names = FALSE)
write.csv(minWide, "minWide.csv", row.names = FALSE)
write.csv(sleepDay, "sleepDay.csv", row.names = FALSE)
write.csv(weightLog, "weightLog.csv", row.names = FALSE)
```

I exported these modified files as a precaution before merging the remaining data sets together. If anything goes wrong, I would be able to start over from this point.


#### Merge Data Sets

```{r, warning=FALSE, message=FALSE}
#join dailyActs & sleepDay datsets by Id & date 
use1 <- full_join(dailyActs, sleepDay, by=join_by(DailyId==SleepDayId, DailyActsDate==SleepDate), keep=TRUE) # keep=TRUE keeps matching column
#remove unnecessary columns 
use1 <- select(use1, -c(SleepDateTime))
```


```{r, warning=FALSE, message=FALSE}
# join daily data and hourly dataset
## merge smaller dataset to larger data set to keep order in Id, Date and Time
use2 <- full_join(use1, hourlyActs, by=join_by(DailyId==HourlyId, DailyActsDate==HourlyDate, SleepTime==Hour), keep=TRUE) 
use2 <- select(use2, -c(ActivityHour))

# clear previous joined files
rm(use1, hourlyActs, dailyActs, sleepDay)
```


```{r, warning=FALSE, message=FALSE, results='hide'}
# join minute datasets together
use3 <- full_join(minActsN, minSleep, by=join_by(MinActsId==MinSleepId, MinuteDates==MinSleepDate, Minutes==SleepMinutes), keep=TRUE)
use3 <- select(use3, -c(DateMinute, MinSleepDateTime))
rm(minActsN, minSleep)
gc()

# combine use2 &  use3
use4 <- full_join(use2, use3, by=join_by(DailyId==MinActsId, DailyActsDate==MinuteDates, Hour==Minutes), keep=TRUE)
rm(use2, use3)
gc()

### will add remaining data sets separately to primary data
# join use4 + hrate
use5 <- full_join(use4, hrate_sec, by=join_by(DailyId==HRateId, DailyActsDate==HRateDate, Minutes==HRateTime), keep=TRUE) 
use5 <- select(use5, -c(HRateDateTime))

# join use4.5 + weight Log
smart_dev_use <- full_join(use5, weightLog, by=join_by(DailyId==WeightId, DailyActsDate==WeightDate, HRateTime==WeightTime), keep=TRUE) 
smart_dev_use <- select(smart_dev_use, -c(WeightDateTime))
rm(use4, use5, hrate_sec, weightLog) #clear environment
gc()
```

I merged all the remaining data sets together except the 'minWide' data set due to its different format. Since I was creating a large data file, I cleared up unnecessary columns and memory in between merging files to avoid a system crash.


#### Export Merged Data Frame

```{r, warning=FALSE, message=FALSE, results='hide'}
# set directory
setwd('C:/Users/Leopoldine/Desktop/Mine/Coding Projects & Portfolio/Bellabeat Case Study/01_tidy_data')

#export file 
write.csv(smart_dev_use, 'smart_dev_use.csv', row.names = FALSE)
```

I made sure to export the merged file as a precaution before making further modifications to the data.


```{r, echo=FALSE, results='hide', include=FALSE}
# clear up memory
gc()
```


### Step 4- Further Modify Combined Dataset

#### Create Primary Id column

```{r}
#new copy in case of error
smartDev <- smart_dev_use  

# Primary Id column == DailyId
# use all except SleepDayId (in range)
smartDev$DailyId <- ifelse(is.na(smartDev$DailyId), smartDev$MinActsId, smartDev$DailyId)
smartDev$DailyId <- ifelse(is.na(smartDev$DailyId), smartDev$HourlyId, smartDev$DailyId)
smartDev$DailyId <- ifelse(is.na(smartDev$DailyId), smartDev$HRateId, smartDev$DailyId)
smartDev$DailyId <- ifelse(is.na(smartDev$DailyId), smartDev$MinSleepId, smartDev$DailyId)
smartDev$DailyId <- ifelse(is.na(smartDev$DailyId), smartDev$WeightId, smartDev$DailyId)

# remove other Id columns
smartDev <- select(smartDev, -c(SleepDayId, HourlyId, MinActsId, HRateId, WeightId, MinSleepId, MinSleepLogId, WeightLogId))

## confirm no more NA values 
sum(is.na(smartDev$DailyId))
```

I turned the 'DailyId' column into the primary 'Id' column. I used the ifelse function to fill the blank spaces with the other Id columns, except for the 'SleepDayId' which was already in range. I confirmed there were no missing values after modifying the column.


```{r, echo=FALSE, results='hide', include=FALSE}
# clear up memory
gc()
```


#### Create Primary Date Column 

```{r}
# Primary Column: DailyActsDate
# use all dates except SleepDate(in range)
smartDev <- smartDev %>% mutate(DailyActsDate=coalesce(DailyActsDate, MinuteDates))
smartDev <- smartDev %>% mutate(DailyActsDate=coalesce(DailyActsDate, HourlyDate))
smartDev <- smartDev %>% mutate(DailyActsDate=coalesce(DailyActsDate, HRateDate))
smartDev <- smartDev %>% mutate(DailyActsDate=coalesce(DailyActsDate, MinSleepDate))
smartDev <- smartDev %>% mutate(DailyActsDate=coalesce(DailyActsDate, WeightDate))

# remove date columns
smartDev <- select(smartDev, -c(SleepDate, HourlyDate, MinuteDates, MinSleepDate, HRateDate, WeightDate))

## confirm no more NA values 
sum(is.na(smartDev$DailyActsDate))
```

I turned the 'DailyActsDate' column into the primary date column, filling the column with the values from the other date columns. I removed the other date columns and confirmed that the primary had no missing values.


#### Create Primary Time Column

```{r}
smartDev <- rename(smartDev, Time=DailyTime)

# merge time columns
smartDev <- smartDev %>% mutate(Time=coalesce(Time, SleepTime))
smartDev <- smartDev %>% mutate(Time=coalesce(Time, Minutes))
smartDev <- smartDev %>% mutate(Time=coalesce(Time, Hour))
smartDev <- smartDev %>% mutate(Time=coalesce(Time, HRateTime))
smartDev <- smartDev %>% mutate(Time=coalesce(Time, SleepMinutes))
smartDev <- smartDev %>% mutate(Time=coalesce(Time, WeightTime))

# remove columns
smartDev <- select(smartDev, -c(SleepTime, Hour, Minutes, SleepMinutes, HRateTime, WeightTime))
```

I used the other time columns to fill in the blanks of the primary 'Time' column. I moved the time column from the end to after the date column for easier readability. 


#### Clean up Modified Data Frame

```{r, results='hide'}
#Rename columns
smartDev <- rename(smartDev, Id=DailyId, Date=DailyActsDate)

# clear environment
rm(smart_dev_use)
gc()
```

I renamed the primary columns to those that are more self-descriptive and appropriate. 


#### View Merged Dataframe

```{r}
str(smartDev)
```


#### Export Final Merged Dataframe

```{r, warning=FALSE, message=FALSE}
setwd('C:/Users/Leopoldine/Desktop/Mine/Coding Projects & Portfolio/Bellabeat Case Study/01_tidy_data')

#export final merged file
write.csv(smartDev, 'smartDev.csv', row.names = FALSE)
```
 
I exported the final merged file as a precaution before manipulating the files further. 


### Step 5 - Transpose Dataframes

#### Pivot Smart Features Dataframe

```{r, warning=FALSE}
# create new dataframe in case of error
smart_d2 <- smartDev

#transpose data 
mod_smart <- 
smart_d2 %>%
  pivot_longer(cols=c(DailyCalories:IsManualReport), 
               names_to = 'SmartFeatures', 
               values_to = 'Values', 
               values_drop_na = TRUE)

#view pivoted data
head(mod_smart) 
```

I decided to transpose (or pivot) the data to have all the smart features in one column and their values in another. As seen above, it cut down the amount of columns and will make it easier to efficiently analyze the data. 


```{r, results='hide', echo=FALSE, include=FALSE}
gc() # clear up storage/memory
```


#### Pivot Minute-Wide Features Dataframe

##### Create Features Column

```{r, warning=FALSE}
## split Cals00-59, Steps00-59, Ints00-59 
mod_minWide <-
minWide %>%
pivot_longer(
    cols=!c(Id, Date, Time),
    names_to = c('Features', '.value'), 
    names_pattern = '(.*)(.\\d)')

#view pivoted data
head(mod_minWide)
```

Since the feature columns were basically formatted as "Feature##", with "##" going from 00 to 59, it made sense to split the column between the word and numbers. I saw there was more to be done after initially transposing the data, especially to cut down on the number of columns.


##### Create Minutes Column

```{r, warning=FALSE}
##create new Minutes column 
mod_mins <-
mod_minWide %>%
pivot_longer(cols = !c(Id, Date, Time, Features),
             names_to = 'Minutes',
             values_to = 'Value') %>%
  rename(Hour=Time) 

#move 'Minutes' after Hour
mod_mins <- mod_mins[,c(1:3,5,4,6)] 
mod_mins$Minutes <- chron(times.=(format(strptime(mod_mins$Minutes, '%M'), '%H:%M:%S')))

#view pivoted data
head(mod_mins)
```

I created a new 'Minutes' column from the '00-59' column range, severely cutting down the number of columns. I also moved the feature values into their own column. I renamed the 'Time' column to 'Hour' since it is more accurate especially with the new column. I ended up changing the data type of the 'Minutes' column and moved it after the 'Hour' column in preparation of merging the two columns.


##### Merge Time Columns

```{r, warning=FALSE}
mod_mins <-
mod_mins %>% 
  mutate(Hour=hours(Hour), Minutes=minutes(Minutes)) %>%
  unite(Time, Hour, Minutes, sep=':') %>%
  mutate(Time=chron(times.=(format(strptime(Time, '%H:%M'), '%H:%M:%S'))))

# view final modified data frame
head(mod_mins)
```

I created a new 'Time' column by merging the 'Hour' and 'Minutes' columns, while also converting it to the 'times' data type. 


```{r, results='hide', include=FALSE}
# Check for NA/Missing Values 
#expect a fair amount as users did not use all features at all times
sum(is.na(mod_smart)) 

##confirm NA locations
apply(X = is.na(mod_smart), MARGIN = 2, FUN = sum)


# should be zero, same as before
sum(is.na(mod_mins))
```

```{r, results='hide', include=FALSE}
#clean environment and memory
rm(minWide, smart_d2, mod_minWide, smartDev)
gc() 
```


#### Export Final Modified Files

```{r, warning=FALSE}
setwd('C:/Users/Leopoldine/Desktop/Mine/Coding Projects & Portfolio/Bellabeat Case Study/01_tidy_data')

write.csv(mod_smart, 'mod_smart.csv', row.names = FALSE)
write.csv(mod_mins, 'mod_mins.csv', row.names = FALSE)
```

I made sure to export both files after all the modifications.


### Step 6 - Analyze Data

#### Descriptive Analysis of Smart Features

```{r}
round(stat.desc(mod_smart), 2)
```


#### Descriptive Analysis of Minute-Wide Features

```{r}
round(stat.desc(mod_mins), 2)
```


#### User Numbers

##### Smart Features User Count

```{r}
mod_smart %>%
  group_by(SmartFeatures)%>%
  summarise(UserCount=length(unique(Id))) %>%
  arrange(desc(UserCount))
```


##### Minute-Wide Features User Count  

```{r}
mod_mins %>%
  group_by(Features)%>%
  summarise(UserCount=length(unique(Id))) %>%
  arrange(desc(UserCount))
```


#### Total Use Count

##### Smart Feature Usage

```{r, warning=FALSE, message=FALSE}
mod_smart %>% count(SmartFeatures, sort = TRUE) %>% rename(TotalCount=n)
```


##### Minute-Wide Features Usage 

```{r}
mod_mins %>% count(Features, sort = TRUE) %>% rename(TotalCount=n)
```


#### Weekday Use Count

##### Total Weekday Smart Feature Usage

```{r, message=FALSE}
mod_smart %>% 
  mutate(Weekday=wday(Date, label=TRUE)) %>%
  group_by(Weekday, SmartFeatures) %>%
  summarise(TotalCount = n()) %>%
  arrange(Weekday, desc(TotalCount)) 
```


##### Total Weekday Minute-Wide Features Usage

```{r, warning=FALSE, message=FALSE}
mod_mins %>% 
  mutate(Weekday=wday(Date, label=TRUE)) %>%
  group_by(Weekday, Features) %>%
  summarise(TotalCount=n()) %>% 
  arrange(Weekday, desc(TotalCount)) 
```


#### Time Frequency 

##### Hourly Frequency of Smart Features

```{r, warning=FALSE, message=FALSE}
mod_smart %>%
  mutate(Hour=hours(Time)) %>%
  group_by(Hour, SmartFeatures) %>%
  summarise(Total_Count=n()) %>%
  arrange(Hour, desc(Total_Count))%>%
  drop_na() 
```


##### Hourly Frequency and Average of Minute-Wide Features

```{r, warning=FALSE, message=FALSE}
mod_mins %>%
  mutate(Hour=hours(Time)) %>%
  group_by(Hour, Features) %>%
  summarise(Total_Count = n(), 
            Average=round(mean(Value), 2)) %>%
  arrange(Hour, desc(Total_Count))
```


##### Minute Frequency of Smart Features 

```{r, warning=FALSE, message=FALSE}
mod_smart %>%
  mutate(Minutes=minutes(Time)) %>%
  group_by(Minutes, SmartFeatures) %>%
  summarise(Total_Count=n()) %>%
  drop_na() %>%
  arrange(Minutes, desc(Total_Count))
```


##### Minute Count & Average of Minute-Wide Features

```{r, warning=FALSE, message=FALSE}
mod_mins %>%
  mutate(Minutes=minutes(Time)) %>%
  group_by(Minutes, Features) %>%
  summarise(Total_Count = n(), 
            Average = round(mean(Value), 2)) %>%
  arrange(Minutes, desc(Total_Count))
```


#### Rank Analysis

##### Top 5 Most Used Features

```{r, warning=FALSE, message=FALSE}
mod_smart %>%
  group_by(SmartFeatures) %>%
  summarise(Total=n()) %>%
  arrange(desc(Total)) %>%
  slice(1:5)
```


##### Top 5 Least Used Features

```{r, warning=FALSE, message=FALSE}
mod_smart %>%
  group_by(SmartFeatures) %>%
  summarise(Total=n()) %>%
  arrange(Total) %>%
  slice(1:5)
```



#### Step 7 - Visualizations

```{r, warning=FALSE, message=FALSE, echo=FALSE}
mod_smart %>%
  group_by(SmartFeatures)%>%
  summarise(User_Count=length(unique(Id))) %>%
  arrange(User_Count) %>%
  ggplot(aes(x=fct_reorder(SmartFeatures,+User_Count), y=User_Count)) +
  geom_col(fill='steelblue') + 
   ggtitle('Total User Count') +
  xlab('Smart Features') +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        plot.title = element_text(hjust = 0.5))
```

Almost all of the smart features were utilized by all 33 users. The ten features that were not are related to three different categories: weight, heart rate and sleep. 


```{r, warning=FALSE, message=FALSE, echo=FALSE}
mod_smart %>%
  group_by(SmartFeatures) %>%
  summarise(Total_Count=n()) %>%
  arrange(desc(Total_Count)) %>%
  slice(1:5) %>%
  ggplot(mapping=aes(x=fct_reorder(SmartFeatures, +Total_Count), y=Total_Count)) +
  geom_col(position = "dodge", fill='maroon') + 
  xlab('Smart Features') +
  ylab('Total Count') +
  ggtitle('Top 5 Most Popular Features') +
  geom_text(aes(label=format(Total_Count, big.mark = ","), y=Total_Count, 
            hjust=ifelse(Total_Count < mean(range(Total_Count)), -0.05, 1.1))) +
  scale_y_continuous(labels = scales::comma) + 
  coord_flip() +
  theme(legend.position = "none", 
        axis.title.y=element_blank(), 
        axis.title.x=element_blank()) 
```

The top 5 most popular smart features among the users possess a time measure (seconds, minutes). The 'HeartRateSec' feature saw nearly twice as much use as the other four smart features. As the time factor of the features increased (from seconds to minutes), the frequency of use decreased. 




```{r, warning=FALSE, message=FALSE, echo=FALSE}
mod_smart %>%
  group_by(SmartFeatures) %>%
  summarise(Total_Count=n()) %>%
  arrange(Total_Count) %>%
  slice(1:5) %>%
  ggplot(mapping=aes(x=fct_reorder(SmartFeatures,+Total_Count), y=Total_Count)) +
  geom_col(stat='identity', position = "dodge", fill='gold3') +
  geom_text(aes(label=Total_Count, y=Total_Count, 
            hjust=ifelse(Total_Count < mean(range(Total_Count)), -0.15, 1.1))) +
  xlab('Smart Features') +
  ylab('Total Count') +
  ggtitle('Top 5 Least Popular Features') +
  coord_flip() +
 theme(legend.position = "none", 
       axis.title.y=element_blank(),
       axis.title.x=element_blank()) 
```

The top 5 least popular features had under 100 uses among users. It should also be noted that these specific features are all related to weight.




```{r, message=FALSE, echo=FALSE}
mod_smart %>% 
  group_by(Date, SmartFeatures) %>%
  summarise(Frequency = n()) %>%
  arrange(Date, desc(Frequency)) %>%
    ggplot(aes(x=Date, y=Frequency)) + 
    geom_point(aes(color=SmartFeatures), position = "jitter") + 
    ggtitle('Daily Frequency of Smart Features') +
    scale_y_continuous(labels = scales::comma) +
    guides(colour = guide_legend(title = NULL, ncol=4)) + 
    theme(legend.position = "bottom", 
          axis.title.x=element_blank(),
          axis.title.y=element_text(size=9),
          legend.text = element_text(size = 8.5), 
          plot.title = element_text(hjust = 0.5)) 
```

There are obviously three major distinct layers of frequency among the smart features over the month long period. Most of the smart features had less than 15,000 frequency of use. The  'HeartRateSec' feature had the highest daily frequency of use with varying amounts.




#### Time Frequency of Smart Features

```{r, warning=FALSE, message=FALSE, echo=FALSE}
library(cowplot)

#area chart: count by hour
hs <- mod_smart %>% 
  mutate(Hour=hours(Time)) %>%
  group_by(Hour) %>%
  summarise(Frequency = n()) %>%
  arrange(Hour) %>%
  drop_na() %>%
    ggplot(aes(x=Hour, y=Frequency)) +
    geom_area(fill='royalblue3', alpha=0.5) +
    geom_line(color='royalblue3', size=1) +
    ggtitle('Hourly Frequency of Smart Features') +
    scale_y_continuous(labels = scales::comma) + 
    theme(plot.title = element_text(size=10, face="bold", hjust = 0.5), 
          axis.title.x=element_blank(),
          axis.title.y = element_blank())

#count by minutes
ms <- mod_smart %>% 
  mutate(Minutes=minutes(Time)) %>%
  group_by(Minutes) %>%
  summarise(Frequency = n()) %>%
  arrange(Minutes) %>%
  drop_na() %>%
    ggplot(aes(x=Minutes, y=Frequency)) +
    geom_area(fill='firebrick', alpha=0.5) +
    geom_line(color='firebrick', size=1) +
    ggtitle('Minute Frequency of Smart Features') +
    scale_y_continuous(labels = scales::comma) + 
    theme(plot.title = element_text(size=10, face="bold", hjust = 0.5), 
          axis.title.x=element_blank(),
          axis.title.y = element_blank())

plot_grid(hs, ms, nrow=2, align="v")
```

The hourly frequency of the smart features was greater than that of the minutes, but not as consistent. The smart features had a frequency of use over 300,000 across the hours. The highest points of use occurred between hours 9 and  19 with over 325,000 uses. The minute frequency had the highest point of use at minute '0' at a little over 225,000. The remaining minutes have a steady frequency over 125,000 uses. 




#### Weekday Analysis of Minute-Wide Features

```{r, message=FALSE, echo=FALSE, warning=FALSE}
mod_mins$Features <- factor(mod_mins$Features, levels=c("Steps", "Calories", "Intensity"))

wm <- mod_mins %>% 
  mutate(Weekday=wday(Date, label=TRUE)) %>%
  group_by(Weekday, Features) %>%
  summarise(Count = n()) %>%
  arrange(Weekday) %>%
    ggplot(aes(x=Weekday, y=Count, fill=Features)) + 
    geom_col() +
    ggtitle('Weekday Frequency of Minute-Wide \n Features') +
    scale_y_continuous(labels = scales::comma) +
    theme(legend.position = "none", 
          plot.title = element_text(face="bold", size=10), 
          axis.title.x=element_blank(),
          axis.title.y=element_blank()) +
    facet_grid(vars(Features))


wam <- mod_mins %>% 
  mutate(Weekday=wday(Date, label=TRUE)) %>%
  group_by(Weekday, Features) %>%
  summarise(Average = mean(Value)) %>%
  arrange(Weekday) %>%
    ggplot(aes(x=Weekday, y=Average, fill=Features)) + 
    geom_bar(stat="identity", position = "dodge") +
    ggtitle('Weekday Average of Minute-Wide Features') +
    scale_y_continuous(labels = scales::comma) +
    theme(plot.title = element_text(face="bold", size=10), 
          axis.title.x=element_blank(), 
          axis.title.y=element_blank(),
          legend.title = element_blank(),
          legend.position = "bottom", 
          legend.text = element_text(size = 7)) 

plot_grid(wm, wam, nrow=1, align = "h")
```

The three 'Minute-Wide' features saw an identical use count above 150,000 throughout the week. The highest point of use was seen on Wednesdays and Thursdays, with over 200,000 count. The features show obvious differences when viewing the Average values, with the 'Steps' feature having the highest value.




#### Time Frequency of Minute-Wide Features

```{r, message=FALSE, echo=FALSE, warning=FALSE}
#stacked area chart---minutes 
mm <- mod_mins %>%
  mutate(Minutes=minutes(Time)) %>%
  group_by(Minutes, Features) %>%
  summarise(Total_Count= n()) %>%
  ggplot(aes(x=Minutes, y=Total_Count, fill=Features)) +
    geom_area(color="white") +
    ggtitle('Minute Frequency of Minute-Wide Features') +
    scale_y_continuous(labels = scales::comma) +
    ylab("Frequency") +
    theme(plot.title = element_text(size=10, face="bold", ), 
          axis.title.x=element_blank(), 
          axis.title.y=element_blank(),
          legend.text = element_text(size=8),
          legend.title = element_blank()) 

am <- mod_mins %>%
  mutate(Minutes=minutes(Time)) %>%
  group_by(Minutes, Features) %>%
  summarise(Average= mean(Value)) %>%
  ggplot(aes(x=Minutes, y=Average, fill=Features)) +
    geom_area(color="white") +
    ggtitle('Minute Averages of Minute-Wide Features') +
    scale_y_continuous(labels = scales::comma) +
    ylab("Average") +
    theme(plot.title = element_text(size=10, face="bold", ), 
          axis.title.x=element_blank(), 
          axis.title.y=element_blank(), 
          legend.text = element_text(size=8),
          legend.title = element_blank()) 


plot_grid(mm, am, nrow=2, align="v")
```

The frequencies of these three features are identical from minute '0' to minute '59', with over 20,000 use count across the minutes. The differences among the features are easily seen when viewing the average values across time, with the 'Steps' feature having the highest value. The steady frequency of use among these features demonstrates their popularity among the users. 




### Top Recommendations

**Drop Least Used Features**

Bellabeat should consider dropping the five least used features. This would allow the company to shift their resources to their more profitable features and better market their products to future customers. 


**Develop 'Heart' Related Features**

Since the 'HeartRateSec' feature saw the most frequent use, the company should develop more heart related features such as blood pressure recorder or a pulse monitor. They can also consider developing a device specifically related to cardio exercises to attract potential customers. 


**Develop 'METS' Related Features**

The company should develop more 'METs' related features as it was among the top five most used features over the month and all 33 users made use of it. This would also expand the available 'Minute-Wide' features (Calories, Intensity, and Steps) used by all users. The increased marketing as a result of the developments would attract more customers.
