---
title: "Bellabeat Case Study"
author: "Leopoldine Mirtil"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Disclaimer

This analysis was made as part of the _Bellabeat Case Study: How Can a Wellness Technology Company Play It Smart?_, offered through the Google Data Analytics Certificate program on Coursera.com. The data is from the [FitBit Fitness Tracker Data](https://www.kaggle.com/datasets/arashnic/fitbit?resource=download) and was originally uploaded to Kaggle.com by the user Möbius for public use. The data covers one month of collection from 4/12/2016 to 5/12/2016 from over 30 consenting users.


## Introduction 

#### Scenario

You are a junior data analyst working on the marketing analyst team at Bellabeat, a high-tech manufacturer of health-focused products for women. Bellabeat is a successful small company, but they have the potential to become a larger player in the global smart device market. Urška Sršen, cofounder and Chief Creative Officer of Bellabeat, believes that analyzing smart device fitness data could help unlock new growth opportunities for the company. You have been asked to focus on one of Bellabeat’s products and analyze smart device data to gain insight into how consumers are using their smart devices. The insights you discover will then help guide marketing strategy for the company. You will present your analysis to the Bellabeat executive team along with your high-level recommendations for Bellabeat’s marketing strategy.


#### Characters

* __Urška__ __Sršen:__ Bellabeat’s cofounder and Chief Creative Officer

* __Sando__ __Mur:__ Mathematician and Bellabeat’s cofounder; key member of the Bellabeat executive team

* __Bellabeat__ __marketing__ __analytics__ __team:__ A team of data analysts responsible for collecting, analyzing, and reporting data that helps guide Bellabeat’s marketing strategy. You joined this team six months ago and have been busy learning about Bellabeat’’s mission and business goals — as well as how you, as a junior data analyst, can help Bellabeat achieve them.


#### Products

* __Bellabeat__ __app:__ The Bellabeat app provides users with health data related to their activity, sleep, stress, menstrual cycle, and mindfulness habits. This data can help users better understand their current habits and make healthy decisions. The Bellabeat app connects to their line of smart wellness products.

* __Leaf:__ Bellabeat’s classic wellness tracker can be worn as a bracelet, necklace, or clip. The Leaf tracker connects to the Bellabeat app to track activity, sleep, and stress.

* __Time:__ This wellness watch combines the timeless look of a classic timepiece with smart technology to track user activity, sleep, and stress. The Time watch connects to the Bellabeat app to provide you with insights into your daily wellness.

* __Spring:__ This is a water bottle that tracks daily water intake using smart technology to ensure that you are appropriately hydrated throughout the day. The Spring bottle connects to the Bellabeat app to track your hydration levels.

* __Bellabeat__ __membership:__ Bellabeat also offers a subscription-based membership program for users. Membership gives users 24/7 access to fully personalized guidance on nutrition, activity, sleep, health and beauty, and mindfulness based on their lifestyle and goals.


### Objective

Sršen asks you to analyze smart device usage data in order to gain insight into how consumers use non-Bellabeat smart devices. She then wants you to select one Bellabeat product to apply these insights to in your presentation.

1. What are some trends in smart device usage?
2. How could these trends apply to Bellabeat customers?
3. How could these trends help influence Bellabeat marketing strategy?


## Get to Work

### Step 1 - Import Data

#### Load Packages

```{r, warning=FALSE, message=FALSE}
library(tidyr)
library(tidyverse)
library(dplyr)
library(knitr)
library(lubridate)
library(chron)
library(pastecs)
library(ggplot2)
library(cowplot)
```

I chose these specific packages to enable: data manipulation, documentation, descriptive statistics and visualization.


```{r, results=FALSE, echo=FALSE}
#remove scientific notation format
options(scipen = 50) 
```


#### Set Directory & Import Data 

```{r, error=FALSE, warning=FALSE}
setwd('C:/Users/Leopoldine/Desktop/Mine/Coding Projects & Portfolio/Bellabeat/00_raw_data')

dailyActs <- read.csv('dailyActivity_merged.csv')
dailyCals <- read.csv('dailyCalories_merged.csv')
dailyInts <- read.csv('dailyIntensities_merged.csv')
dailySteps <- read.csv('dailySteps_merged.csv')
hrate_sec <- read.csv('heartrate_seconds_merged.csv')
hrCals <- read.csv('hourlyCalories_merged.csv')
hrInts <- read.csv('hourlyIntensities_merged.csv')
hrSteps <- read.csv('hourlySteps_merged.csv')
minCalsN <- read.csv('minuteCaloriesNarrow_merged.csv')  
minCalsW <- read.csv('minuteCaloriesWide_merged.csv')
minIntsN <- read.csv('minuteIntensitiesNarrow_merged.csv')
minIntsW <- read.csv('minuteIntensitiesWide_merged.csv')
minMETsN <- read.csv('minuteMETsNarrow_merged.csv')
minSleep <- read.csv('minuteSleep_merged.csv')
minStepsN <- read.csv('minuteStepsNarrow_merged.csv')
minStepsW <- read.csv('minuteStepsWide_merged.csv')
sleepDay <- read.csv('sleepDay_merged.csv')
weightLog <- read.csv('weightLogInfo_merged.csv')
```

I set the directory first to make it easier to import all the data files without having to include the complete file path every time.


### Step 2 - Clean Data

#### Review Daily Dataframes

I noticed that the dailyActs file had columns that were similar to those in the other daily data files. First, I decided to confirm if these columns were identical, and if so, remove those files.


##### Compare Daily Dataframes

```{r}
##dailyActs vs dailyCals
# TRUE = equal, False=not equal
all.equal(dailyActs$Calories, dailyCals$Calories) 
all.equal(dailyActs$Id, dailyCals$Id)
all.equal(dailyActs$ActivityDate, dailyCals$ActivityDay)
```


```{r}
##dailyActs vs dailySteps
all.equal(dailyActs$TotalSteps, dailySteps$StepTotal) 
all.equal(dailyActs$Id, dailySteps$Id)
all.equal(dailyActs$ActivityDate, dailySteps$ActivityDay)
```


```{r}
#dailyActs vs dailyInts
all.equal(dailyActs$Id, dailyInts$Id)
all.equal(dailyActs$ActivityDate, dailyInts$ActivityDay)
```


```{r}
#dailyActs vs dailyInts
##convert columns into data tables for comparison
d_Acts <- data.table::setDT(dailyActs[7:14])
d_Ints <- data.table::setDT(dailyInts[3:10])

all.equal(d_Acts, d_Ints, ignore.col.order = TRUE)
```

I needed to convert the remaining columns in dailyActs & dailyInts data frames into data tables (due to their different column orders) in order to perform a comparison.


```{r}
#compare similar columns 'TotalDistance' vs 'TrackerDistance'
tod <- data.table::setDT(dailyActs[5])
trd <- data.table::setDT(dailyActs[6])

all.equal(tod, trd)

#remove a column
dailyActs <- select(dailyActs, -c(TotalDistance))
```

I removed the TotalDistance column after confirming it was identical to the TrackerDistance column.  I kept this column because it has the more illustrative name of the two. 


```{r, results='hide', include=FALSE}
#clean environment
#remove unneeded datasets
rm(tod, trd, d_Acts, d_Ints, dailyCals, dailyInts, dailySteps)
gc()
```


##### Modify Daily Dataframe

```{r}
dailyActs$ActivityDate <- as.Date(dailyActs$ActivityDate, '%m/%d/%Y') 
```

After removing the unnecessary 'daily' files, I changed the data type of the ActivityDate column as I will use it as a key point when merging the data sets later on. 


#### Review Hourly Dataframes

##### Compare Hourly Dataframes

```{r}
#check Id columns
all.equal(hrCals$Id, hrInts$Id)
all.equal(hrSteps$Id, hrInts$Id)

#check Activity columns
all.equal(hrCals$ActivityHour, hrInts$ActivityHour)
all.equal(hrSteps$ActivityHour, hrInts$ActivityHour)
```


##### Merge Hourly Dataframes

```{r}
hourlyActs <- bind_cols(hrCals, hrSteps[3], hrInts[3:4])

# view data
str(hourlyActs)
```


```{r, results='hide', include=FALSE}
#clean environment and remove unneeded datasets
rm(hrCals, hrInts, hrSteps)

#clear up memory
gc() 
```


#### Review Minute Dataframes

##### Compare Minute Dataframes

```{r}
# check Id&ActivityMinute columns 
all.equal(minCalsN[1:2],minStepsN[1:2])
all.equal(minIntsN[1:2],minMETsN[1:2])
all.equal(minCalsN[1:2],minMETsN[1:2])

all.equal(minCalsW[1:2],minStepsW[1:2])
all.equal(minIntsW[1:2],minCalsW[1:2])
```


##### Merge Minute-Narrow Dataframes

```{r}
minActs <- bind_cols(minCalsN, minStepsN[3], minIntsN[3], minMETsN[3])

str(minActs) 
```


##### Merge Minute-Wide Dataframes

```{r}
minWide <- bind_cols(minCalsW, minIntsW[,3:62], minStepsW[,3:62]) 
```


```{r, results='hide', include=FALSE}
#clean environment
rm(minCalsN, minStepsN, minIntsN, minMETsN, minCalsW, minIntsW, minStepsW)
gc()
```


### Step 3 - Modify Data

#### Rename Columns

```{r}
dailyActs <- rename(dailyActs, DailyActsDate=ActivityDate, DailyId=Id)
sleepDay <- rename(sleepDay, DateTime=SleepDay, SleepDayId=Id)
hourlyActs <- rename(hourlyActs, DateTime=ActivityHour, TotalSteps=StepTotal, HourlyId=Id)
hrate_sec <- rename(hrate_sec, HRateId=Id, DateTime=Time, HeartRate=Value)
minActs <- rename(minActs, MinActsId=Id, DateTime=ActivityMinute)
minSleep <- rename(minSleep, MinSleepId=Id, DateTime=date, Sleep=value)
weightLog <- rename(weightLog, WeightId=Id, DateTime=Date)
minWide <- rename(minWide, DateTime=ActivityHour)
```

I changed the names of these columns to be more identifiable when modifying and merging the data.


#### Remove Columns

```{r}
minSleep <- select(minSleep, -c(logId))
weightLog <- select(weightLog, -c(LogId))
```

I removed these columns because they would be redundant as the Id column exists and would allow easier analysis.


#### Convert Logical Values to Integers 

```{r}
weightLog$IsManualReport <- as.integer(as.logical(weightLog$IsManualReport))
```

I converted the logical values of the column into integers where TRUE=1 and FALSE=0 to enable easier analysis of this feature.


#### Add New Columns

##### Add Date-Only Columns

```{r}
hourlyActs$HourlyDate <- as.Date(hourlyActs$DateTime, '%m/%d/%Y')
minActs$MinuteDate <- as.Date(minActs$DateTime, '%m/%d/%Y')
minSleep$MinSleepDate <- as.Date(minSleep$DateTime, '%m/%d/%Y')
sleepDay$SleepDate <- as.Date(sleepDay$DateTime, '%m/%d/%Y')
weightLog$WeightDate <- as.Date(weightLog$DateTime, '%m/%d/%Y')
hrate_sec$HRateDate <- as.Date(hrate_sec$DateTime, '%m/%d/%Y')
minWide$Date <- as.Date(minWide$DateTime, '%m/%d/%Y %r')
```


##### Add Time-Only Columns 

```{r, warning=FALSE}
hourlyActs$Hour <- chron(times.=(format(strptime(hourlyActs$DateTime, '%m/%d/%Y %r'), '%H:%M:%S')))
minActs$Minutes<- chron(times.=(format(strptime(minActs$DateTime, '%m/%d/%Y %r'), '%H:%M:%S')))
minSleep$SleepMinutes <- chron(times.=(format(strptime(minSleep$DateTime, '%m/%d/%Y %r'), '%H:%M:%S')))
sleepDay$SleepTime <- chron(times.=(format(strptime(sleepDay$DateTime, '%m/%d/%Y %r'), '%H:%M:%S')))
weightLog$WeightTime <- chron(times.=(format(strptime(weightLog$DateTime, '%m/%d/%Y %r'), '%H:%M:%S')))
hrate_sec$HRateTime <- chron(times.=(format(strptime(hrate_sec$DateTime, '%m/%d/%Y %r'), '%H:%M:%S')))
minWide$Time <- chron(times.=(format(strptime(minWide$DateTime, '%m/%d/%Y %r'), '%H:%M:%S')))
dailyActs$Time <- chron(times.=(as.numeric(NA))) 
```

I added separate date and time columns to the data to make it easier to identify patterns or trends. I added a blank time column to the dailyActs data frame because it will be a one of the primary columns when merging the files together. 


#### Remove DateTime Columns

```{r}
hourlyActs <- select(hourlyActs, -c(DateTime))
minSleep <- select(minSleep, -c(DateTime))
weightLog <- select(weightLog, -c(DateTime))
minActs <-select(minActs, -c(DateTime))
sleepDay <- select(sleepDay, -c(DateTime))
hrate_sec <- select(hrate_sec, -c(DateTime))
minWide <- select(minWide, -c(DateTime))
```


#### Reorganize Columns

```{r}
hourlyActs <- hourlyActs[, c(1, 6, 7, 2:5)]
minActs <- minActs[, c(1, 6, 7, 2:5)]
minSleep <- minSleep[, c(1, 3, 4, 2)]
sleepDay <- sleepDay[, c(1, 5, 6, 2:4)]
weightLog <- weightLog[, c(1, 7, 8, 2:6)]
hrate_sec <- hrate_sec[, c(1, 3, 4, 2)]
minWide <- minWide[, c(1, 182, 183, 2:181)]
dailyActs <- dailyActs[, c(1, 2, 15, 14, 3:13)] 
```

I moved the new date and time columns for readability and to make it easier to merge the data sets all together. I also took the time to move the Calories column from the end of the dailyActs data frame.


#### Add Suffix to Columns

```{r}
library(datawizard)

dailyActs <- data_addsuffix(dailyActs,"_Daily", Calories:SedentaryMinutes)
minActs <- data_addsuffix(minActs,"_Minute", Calories:METs)
minSleep <- data_addsuffix(minSleep ,"_Minute", Sleep)
sleepDay <- data_addsuffix(sleepDay,"_Daily", TotalSleepRecords:TotalTimeInBed)
hourlyActs <- data_addsuffix(hourlyActs,"_Hourly", Calories:AverageIntensity)
weightLog <- data_addsuffix(weightLog,"_Daily", WeightKg:IsManualReport)
hrate_sec <- data_addsuffix(hrate_sec,"_Seconds", HeartRate)
```

Since the data was recorded at different time intervals, I felt that it would be important to make note of this when analyzing the data later on. To that end, I added a suffix to the feature columns to mark what time interval they were recorded in. Later on, I can split the interval into its own column.


#### Confirm Date Ranges

The data was collected over a one month period from 4/12/2016 to 5/12/2016. I wanted to verify that all the dates of the files are within range.

```{r}
range(dailyActs$DailyActsDate)

range(hourlyActs$HourlyDate) 

range(minActs$MinuteDate)

range(minSleep$MinSleepDate) 

range(sleepDay$SleepDate) 

range(weightLog$WeightDate) 

range(hrate_sec$HRateDate)  

range(minWide$Date) 
```

The only data sets containing out of range dates are minSleep (4/11/2016 - 5/12/2016) and minWide (4/13/2016- 5/13/2016). 


#### Remove Out of Range Dates

```{r}
minSleep <- minSleep[minSleep[['MinSleepDate']]>='2016-04-12', ] 

minWide <- minWide[minWide[['Date']]<='2016-05-12', ]

# confirm date range
range(minSleep$MinSleepDate)
range(minWide$Date)
```


```{r, results='hide', include=FALSE}
#calculate missing values
sum(is.na(dailyActs)) 
sum(is.na(hourlyActs))
sum(is.na(minActs)) 
sum(is.na(minSleep))
sum(is.na(sleepDay)) 
sum(is.na(weightLog)) 
sum(is.na(hrate_sec))
sum(is.na(minWide)) 
```


```{r, results='hide', include=FALSE}
# examine 'weightLog' data for location of all NA vals
apply(X = is.na(weightLog), MARGIN = 2, FUN = sum)
#won't remove NA values to avoid possibility of skewing results
```


#### Export Modified Dataframes

```{r, warning=FALSE}
setwd('C:/Users/Leopoldine/Desktop/Mine/Coding Projects & Portfolio/Bellabeat/01_tidy_data')

write.csv(dailyActs, "dailyActs.csv", row.names = FALSE)
write.csv(hourlyActs, "hourlyActs.csv", row.names = FALSE)
write.csv(hrate_sec, "hrate_sec.csv", row.names = FALSE)
write.csv(minActs, "minActs.csv", row.names = FALSE)
write.csv(minSleep, "minSleep.csv", row.names = FALSE)
write.csv(sleepDay, "sleepDay.csv", row.names = FALSE)
write.csv(weightLog, "weightLog.csv", row.names = FALSE)
write.csv(minWide, "minWide.csv", row.names = FALSE)
```

I exported these modified files as a precaution before merging the remaining data sets together. If anything goes wrong, I would be able to start over from this point.


#### Merge Dataframes

```{r, warning=FALSE, message=FALSE}
#join dailyActs & sleepDay datsets by Id & date as keys
daily <- full_join(dailyActs, sleepDay, by=join_by(DailyId==SleepDayId, DailyActsDate==SleepDate), keep=TRUE) # keep=TRUE keeps matching column

gc()
```


```{r, warning=FALSE, message=FALSE}
# join daily data and hourly dataset
## merge smaller dataset to larger data set to keep order in Id, Date and Time
daily_hour <- full_join(daily, hourlyActs, by=join_by(DailyId==HourlyId, DailyActsDate==HourlyDate, SleepTime==Hour), keep=TRUE) 

# clear previous joined files
rm(daily, hourlyActs, dailyActs, sleepDay)
gc()
```


```{r, warning=FALSE, message=FALSE, results='hide'}
# join minute datasets together
minutes <- full_join(minActs, minSleep, by=join_by(MinActsId==MinSleepId, MinuteDate==MinSleepDate, Minutes==SleepMinutes), keep=TRUE)

rm(minActs, minSleep)
gc()

# combine daily_hour &  minutes; use hour as next longest available time format
d_hr_min <- full_join(daily_hour, minutes, by=join_by(DailyId==MinActsId, DailyActsDate==MinuteDate, Hour==Minutes), keep=TRUE)
rm(daily_hour, minutes)
gc()

# join d_hr_min + hrate_sec
smarties <- full_join(d_hr_min, hrate_sec, by=join_by(DailyId==HRateId, DailyActsDate==HRateDate, Minutes==HRateTime), keep=TRUE) 

# join smarties + weight Log
smart <- full_join(smarties, weightLog, by=join_by(DailyId==WeightId, DailyActsDate==WeightDate, HRateTime==WeightTime), keep=TRUE) 

#clear environment
rm(smarties, d_hr_min, hrate_sec, weightLog) 
gc()
```

I merged all the remaining files together, with the exception of the minWide (due to its different format). Since I was creating a large data file, I cleared the environment and memory in between merging to avoid a system crash.


#### Export Merged Dataframe

```{r, warning=FALSE, message=FALSE, results='hide'}
# set directory
setwd('C:/Users/Leopoldine/Desktop/Mine/Coding Projects & Portfolio/Bellabeat/01_tidy_data')

#export file 
write.csv(smart, 'smart.csv', row.names = FALSE)
```


```{r, echo=FALSE, results='hide', include=FALSE}
# clear up memory
gc()
```


### Step 4 - Further Modify Combined Dataset

#### Create Primary Id Column

```{r}
#new copy in case of error
smartDev <- smart 

smartDev$DailyId <- ifelse(is.na(smartDev$DailyId), smartDev$MinActsId, smartDev$DailyId)
smartDev$DailyId <- ifelse(is.na(smartDev$DailyId), smartDev$HourlyId, smartDev$DailyId)
smartDev$DailyId <- ifelse(is.na(smartDev$DailyId), smartDev$HRateId, smartDev$DailyId)
smartDev$DailyId <- ifelse(is.na(smartDev$DailyId), smartDev$MinSleepId, smartDev$DailyId)
smartDev$DailyId <- ifelse(is.na(smartDev$DailyId), smartDev$WeightId, smartDev$DailyId)


## confirm no more NA values 
sum(is.na(smartDev$DailyId))

# remove other Id columns
smartDev <- select(smartDev, -c(SleepDayId, HourlyId, MinActsId, HRateId, WeightId, MinSleepId))
```

I made the DailyId column the key Id column and filled in the blanks with the other Id columns(except SleepId which was in range). I confirmed there were no missing values once completed, then removed the other unnecessary Id columns.


```{r, echo=FALSE, results='hide', include=FALSE}
# clear up memory
gc()
```


#### Create Primary Date Column 

```{r}
smartDev <- 
  smartDev %>% 
    mutate(DailyActsDate=coalesce(DailyActsDate, SleepDate)) %>%
    mutate(DailyActsDate=coalesce(DailyActsDate, MinuteDate)) %>%
    mutate(DailyActsDate=coalesce(DailyActsDate, HourlyDate)) %>%
    mutate(DailyActsDate=coalesce(DailyActsDate, HRateDate)) %>%
    mutate(DailyActsDate=coalesce(DailyActsDate, MinSleepDate)) %>%
    mutate(DailyActsDate=coalesce(DailyActsDate, WeightDate))

## confirm no more NA values 
sum(is.na(smartDev$DailyActsDate))

# remove other date columns
smartDev <- select(smartDev, -c(SleepDate, HourlyDate, MinuteDate, MinSleepDate, HRateDate, WeightDate))
```

I turned the DailyActsDate column into the primary date column, filling the blanks with the values from the other date columns. I removed the other columns after confirming that there were no missing values in the primary column.


#### Create Primary Time Column

```{r}
smartDev <- 
  smartDev %>% 
    mutate(Time=coalesce(Time, SleepTime)) %>% 
    mutate(Time=coalesce(Time, Minutes)) %>% 
    mutate(Time=coalesce(Time, Hour)) %>% 
    mutate(Time=coalesce(Time, HRateTime)) %>% 
    mutate(Time=coalesce(Time, SleepMinutes)) %>% 
    mutate(Time=coalesce(Time, WeightTime))

# remove columns
smartDev <- select(smartDev, -c(SleepTime, Hour, Minutes, SleepMinutes, HRateTime, WeightTime))
```

I used the other time columns to fill in the blanks of the primary Time column. I moved the time column from the end to after the date column for easier readability. 


#### Clean up Modified Dataframe

```{r, results='hide'}
#Rename columns
smartDev <- rename(smartDev, Id=DailyId, Date=DailyActsDate)

# clear environment
rm(smart)
gc()
```

I renamed the primary columns to those that are more self-descriptive and appropriate. 


#### View Merged Dataframe

```{r}
str(smartDev)
```


#### Export Final Merged Dataframe

```{r, warning=FALSE, message=FALSE}
setwd('C:/Users/Leopoldine/Desktop/Mine/Coding Projects & Portfolio/Bellabeat/01_tidy_data')

#export final merged file
write.csv(smartDev, 'smartDev.csv', row.names = FALSE)
```
 
I exported the final merged file as a precaution before manipulating the files further. 


### Step 5 - Final Manipulation of Dataframes

#### Modify Smart Features

##### Pivot Smart Features Dataframe

```{r, warning=FALSE}
# create new dataframe in case of error
smart_d2 <- smartDev

#transpose data 
mod_smart <- 
smart_d2 %>%
  pivot_longer(cols=c(Calories_Daily:IsManualReport_Daily), 
               names_to = 'SmartFeatures', 
               values_to = 'Values', 
               values_drop_na = TRUE)

#view pivoted data
head(mod_smart) 
```

I decided to transpose (or pivot) the data to have all the smart features in one column and their values in another.  


```{r, results='hide', echo=FALSE, include=FALSE}
gc() # clear up storage/memory
```


##### Create Interval Column

```{r}
mod_smart <-
mod_smart %>%
  separate_wider_delim(SmartFeatures, "_", names=c('SmartFeatures', 'Interval'))

head(mod_smart)
```

I created this new column by splitting the 'SmartFeaures' column to remove the interval suffix I had added earlier. 


##### Add Category Column

```{r}
mod_smart$Category <- with(mod_smart, ifelse(SmartFeatures == "Calories", 'Calories',
                            ifelse(SmartFeatures == "METs", 'METs',
                            ifelse(SmartFeatures == "Sleep" |SmartFeatures =="TotalSleepRecords"
                                   |SmartFeatures =="TotalMinutesAsleep"
                                   |SmartFeatures =="TotalTimeInBed", 'Sleep', 
                            ifelse(SmartFeatures == "Steps" |SmartFeatures =="TotalSteps", 'Steps', 
                            ifelse(SmartFeatures == "WeightKg"|SmartFeatures =="WeightPounds"
                                   |SmartFeatures =="Fat"|SmartFeatures =="BMI"
                                   |SmartFeatures =="IsManualReport", 'Weight', 
                            ifelse(SmartFeatures =="HeartRate", 'Heart Rate', 'Intensity')))))))
```

I added the Category column in order to group related features which would allow easier analysis of the features. 


##### Reorganize Columns

```{r}
mod_smart <- mod_smart[, c(1:4, 6, 7, 5)] 
head(mod_smart)
```



```{r, results='hide', echo=FALSE, include=FALSE}
#clear up storage/memory
gc() 

#clear enivronment
rm(smart_d2, smartDev)
```


#### Modify Minute-Wide Features Dataframe

##### Pivot Minute-Wide Features Dataframe

```{r, warning=FALSE}
## split Cals00-59, Steps00-59, Ints00-59 
mod_minWide <-
minWide %>%
pivot_longer(
    cols=!c(Id, Date, Time),
    names_to = c('Features', '.value'), 
    names_pattern = '(.*)(.\\d)')

#view pivoted data
head(mod_minWide)
```

Since the columns were originally in the "Feature##" format, with "##" covering minutes '00-59', it made sense to split the columns between the words and numbers. 


##### Create Minutes Column

```{r, warning=FALSE}
mod_minWide <-
mod_minWide %>%
pivot_longer(cols = !c(Id, Date, Time, Features),
             names_to = 'Minutes',
             values_to = 'Value')

#move 'Minutes' after Time(contains 'hours')
mod_minWide <- mod_minWide[,c(1:3,5,4,6)] 
mod_minWide$Minutes <- chron(times.=(format(strptime(mod_minWide$Minutes, '%M'), '%H:%M:%S')))

#view pivoted data
head(mod_minWide)
```

I created the new Minutes column from the '00-59' column range. I also moved the feature values into their own column. I also took the time to change the data type of the Minutes column and move it in preparation of merging both time columns.


##### Merge Time Columns

```{r, warning=FALSE}
mod_minWide <-
mod_minWide %>% 
  mutate(Hour=hours(Time), Minutes=minutes(Minutes)) %>%
  unite(Time, Hour, Minutes, sep=':') %>%
  mutate(Time=chron(times.=(format(strptime(Time, '%H:%M'), '%H:%M:%S'))))

# view final modified data frame
head(mod_minWide)
```


```{r, results='hide', echo=FALSE, include=FALSE}
# clear up storage/memory and environment
gc() 
rm(minWide)
```


#### Export Final Modified Files

```{r, warning=FALSE}
setwd('C:/Users/Leopoldine/Desktop/Mine/Coding Projects & Portfolio/Bellabeat/01_tidy_data')

write.csv(mod_smart, 'mod_smart.csv', row.names = FALSE)
write.csv(mod_minWide, 'mod_minWide.csv', row.names = FALSE)
```

I made sure to export both files after completing all of the data modifications.


### Step 5 - Analyze Data

#### Descriptive Analysis of Smart Features

```{r}
round(stat.desc(mod_smart), 2)
```


#### Descriptive Analysis of Minute-Wide Features

```{r}
round(stat.desc(mod_minWide), 2)
```


#### User Numbers

##### Smart Features User Count

```{r}
mod_smart %>%
  group_by(SmartFeatures)%>%
  summarise(UserCount=length(unique(Id))) %>%
  arrange(desc(UserCount)) %>%
  print(n=27)
```


##### Minute-Wide User Count

```{r}
mod_minWide %>%
  group_by(Features)%>%
  summarise(UserCount=length(unique(Id))) %>%
  arrange(desc(UserCount)) 
```


#### Total Use Count

##### Feature Category Usage

```{r, warning=FALSE, message=FALSE}
mod_smart %>%
  group_by(Category) %>%
  summarise(Total=n()) %>%
  arrange(desc(Total)) 
```


##### Smart Features Usage by Category

```{r, warning=FALSE, message=FALSE}
mod_smart %>%
  group_by(Category, SmartFeatures)%>%
  summarise(Count=n()) %>%
  arrange(Category, desc(Count)) %>%
  print(n=27)
```


##### Minute-Wide Features Usage 

```{r}
mod_minWide %>% count(Features, sort = TRUE) %>% rename(TotalCount=n)
```


##### Total Count by Interval 

```{r, warning=FALSE, message=FALSE}
mod_smart %>%
  group_by(Interval) %>%
  summarise(Count=n()) %>%
  arrange(desc(Count))
```


##### Total Count by Category and Interval 

```{r, warning=FALSE, message=FALSE}
mod_smart %>%
  group_by(Interval, Category) %>%
  summarise(Count=n()) %>%
  arrange(Interval, desc(Count))
```

#### Daily Frequency 

##### Daily Frequency of Feature Categories

```{r, message=FALSE}
mod_smart %>% 
  group_by(Date, Category) %>%
  summarise(TotalCount = n()) %>%
  arrange(Date, desc(TotalCount)) 
```


##### Frequency of Daily-Tracked Features

```{r, message=FALSE}
mod_smart %>% 
  filter(Interval=="Daily") %>%
  group_by(Date, SmartFeatures) %>%
  summarise(TotalCount = n()) %>%
  arrange(Date, desc(TotalCount)) 
```


#### Weekday Frequency

##### Weekday Frequency of Feature Categories

```{r, message=FALSE}
mod_smart %>% 
  mutate(Weekday=wday(Date, label=TRUE)) %>%
  group_by(Weekday, Category) %>%
  summarise(TotalCount = n()) %>%
  arrange(Weekday, desc(TotalCount)) 
```


##### Weekday Frequency of Smart Features

```{r, message=FALSE}
mod_smart %>% 
  mutate(Weekday=wday(Date, label=TRUE)) %>%
  group_by(Weekday, SmartFeatures) %>%
  summarise(TotalCount = n()) %>%
  arrange(Weekday, desc(TotalCount)) 
```


##### Weekday Frequency of Minute-Wide Features 

```{r, warning=FALSE, message=FALSE}
mod_minWide %>% 
  mutate(Weekday=wday(Date, label=TRUE)) %>%
  group_by(Weekday, Features) %>%
  summarise(TotalCount=n()) %>% 
  arrange(Weekday, desc(TotalCount)) 
```


#### Time Frequency 

##### Hourly Frequency of Smart Features

```{r, warning=FALSE, message=FALSE}
mod_smart %>%
  filter(Interval=="Hourly") %>%
  mutate(Hour=hours(Time)) %>%
  group_by(Hour, SmartFeatures) %>%
  summarise(Total_Count=n()) %>%
  arrange(Hour, desc(Total_Count))
```


##### Minute Frequency of Smart Features 

```{r, warning=FALSE, message=FALSE}
mod_smart %>%
  filter(Interval=="Minute") %>%
  mutate(Minutes=minutes(Time)) %>%
  group_by(Minutes, SmartFeatures) %>%
  summarise(Total_Count=n()) %>%
  arrange(Minutes, desc(Total_Count))
```


##### Minute Frequency & Average Value of Minute-Wide Features

```{r, warning=FALSE, message=FALSE}
mod_minWide %>%
  mutate(Minutes=minutes(Time)) %>%
  group_by(Minutes, Features) %>%
  summarise(Total_Count = n(), 
            Average = round(mean(Value), 2)) %>%
  arrange(Minutes, desc(Total_Count))
```


#### Rank Analysis

##### Top 5 Most Used Features

```{r, warning=FALSE, message=FALSE}
mod_smart %>%
  group_by(SmartFeatures) %>%
  summarise(Total=n()) %>%
  arrange(desc(Total)) %>%
  slice(1:5)
```

##### Bottom 5 Least Used Features

```{r, warning=FALSE, message=FALSE}
mod_smart %>%
  group_by(SmartFeatures) %>%
  summarise(Total=n()) %>%
  arrange(Total) %>%
  slice(1:5)
```

### Step 6 - Visualizations

```{r, warning=FALSE, message=FALSE, echo=FALSE}
mod_smart %>%
  group_by(Category) %>%
  summarise(Total=n()) %>%
  mutate(percentage=round(Total*100/sum(Total), 3)) %>%
  arrange(desc(Category)) %>%
  mutate(y_pos = cumsum(percentage)-0.5*percentage) %>%
  ggplot(aes(x=2, y=percentage, fill=Category))+
    geom_bar(stat="identity", color="white") +
    ggtitle('Total Use by Feature Category') +
    coord_polar(theta = "y", start=0)+
    geom_text(aes(y=y_pos, label=paste0(Category, sep="\n", percentage, "%")), color="black", size=3)+
    theme_void() +
    theme(legend.position = "none",
          plot.title = element_text(hjust = 0.5)) +
    xlim(0.6,2.6)
```

The Heart Rate category saw the most use during the month-long period. The Weight and Sleep categories saw less than 5% of use combined during the period. Since the categories reflect their related smart features, we can assume that the heart rate feature saw the most use during the period while both weight-related and sleep-related features saw the least.


```{r, warning=FALSE, message=FALSE, echo=FALSE}
mod_smart %>%
  group_by(SmartFeatures) %>%
  summarise(Total_Count=n()) %>%
  arrange(desc(Total_Count)) %>%
  slice(1:5) %>%
  ggplot(mapping=aes(x=fct_reorder(SmartFeatures, +Total_Count), y=Total_Count)) +
  geom_col(position = "dodge", fill='maroon') + 
  xlab('Smart Features') +
  ylab('Total Count') +
  ggtitle('Top 5 Most Used Features') +
  geom_text(aes(label=format(Total_Count, big.mark = ","), y=Total_Count, 
            hjust=ifelse(Total_Count < mean(range(Total_Count)), -0.05, 1.1))) +
  scale_y_continuous(labels = scales::comma) + 
  coord_flip() +
  theme(legend.position = "none", 
        axis.title.y=element_blank(), 
        axis.title.x=element_blank()) 
```

The top 5 most used features recorded over 1,300,000 use count over the course of the month. The heart rate feature saw nearly twice as much use as the other four, demonstrating its preference among users.


```{r, warning=FALSE, message=FALSE, echo=FALSE}
mod_smart %>%
  group_by(SmartFeatures) %>%
  summarise(Total_Count=n()) %>%
  arrange(Total_Count) %>%
  slice(1:5) %>%
  ggplot(mapping=aes(x=fct_reorder(SmartFeatures,+Total_Count), y=Total_Count)) +
  geom_col(stat='identity', position = "dodge", fill='gold3') +
  geom_text(aes(label=Total_Count, y=Total_Count, 
            hjust=ifelse(Total_Count < mean(range(Total_Count)), -0.15, 1.1))) +
  xlab('Smart Features') +
  ylab('Total Count') +
  ggtitle('Bottom 5 Least Used Features') +
  coord_flip() +
 theme(legend.position = "none", 
       axis.title.y=element_blank(),
       axis.title.x=element_blank()) 
```

The bottom five least used features recorded under 100 use count during the month, indicating low engagement by the users. Important to note that all five of these features are related to weight.


```{r, warning=FALSE, message=FALSE, echo=FALSE}
mod_smart %>%
  group_by(Interval, Date) %>%
  summarise(UseCount=n()) %>%
  arrange(desc(UseCount)) %>%
  ggplot(aes(x=Date, y=UseCount, color=Interval, group=Interval)) +
    geom_point() + 
    geom_line() +
    ggtitle("Daily Frequency by Time Interval") +
    scale_color_brewer(palette = "Dark2") +
    scale_y_continuous(labels = scales::comma) + 
    theme(axis.title.y=element_blank())
```

The highest frequency of use was recorded by minutes with over 150,000 uses steadily throughout most of the month. This shows that users greatly preferred tracking their activities first by minute, then by seconds compared to the other time intervals.


```{r, warning=FALSE, message=FALSE, echo=FALSE}
mod_smart %>%
  group_by(Interval, Category) %>%
  summarise(Total=n()) %>%
  mutate(percentage=round(Total/sum(Total), 4)) %>%
  arrange(Interval, percentage) %>%
  ggplot(aes(x=Interval, y=percentage, fill=Category)) +
    geom_col(stat="identity") +
    geom_text(aes(label=if_else(percentage > 0.05, paste0(sprintf("%1.2f", percentage*100),"%"), ""),
                  width=0.75), position=position_stack(vjust=0.5), size=4, color="white") +
    scale_color_brewer(palette = "paired") +
    scale_y_continuous(labels = scales::percent) + 
    ggtitle('Total Category Usage by Time Interval') +
    theme(axis.title.x=element_blank(),
          axis.title.y=element_blank()) 
```

The Intensity category saw the most daily use, while the Weight category saw the least. The Hourly interval recorded use from 3 feature categories, with the Intensity category made up half of the total use. The Minute interval recorded use from 5 categories, with the Sleep category recording the less than 4% of use. The only feature tracked during seconds was heart rate. With the exception of Seconds, as the time interval increased, so did the number of categories (and related features) use among users.



```{r, message=FALSE, echo=FALSE, warning=FALSE}
mod_smart %>% 
  filter(Interval=="Daily") %>%
  group_by(Date, SmartFeatures) %>%
  summarise(Frequency = n()) %>%
  arrange(Date, desc(Frequency)) %>%
    ggplot(aes(x=Date, y=Frequency)) + 
    geom_point(aes(color=SmartFeatures), position = "jitter") + 
    ggtitle('Frequency of Daily-Tracked Smart Features') +
    scale_y_continuous(labels = scales::comma) +
    guides(colour = guide_legend(title = NULL, ncol=4)) + 
    theme(legend.position = "bottom",
      axis.title.x=element_blank(),
          axis.title.y=element_blank(),
          legend.text = element_text(size = 8.5), 
          plot.title = element_text(hjust = 0.5)) 
```

There are three distinct layers of frequency among the smart features over the month. The Calories features and all the intensity-related features had the highest daily use out of all of the daily-tracked features. The bottom layer was made up of all of the weight-related features and saw a daily use of less than 5.


```{r, message=FALSE, echo=FALSE, warning=FALSE}
hf <- mod_smart %>% 
  filter(Interval=="Hourly") %>%
  mutate(Hour=hours(Time)) %>%
  group_by(Hour, SmartFeatures) %>%
  summarise(Frequency = n()) %>%
  arrange(Hour) %>%
    ggplot(aes(x=Hour, y=Frequency, fill=SmartFeatures, group=SmartFeatures)) +
    geom_bar(stat="identity") + 
    ggtitle('Frequency of Hourly-Tracked Smart Features') +
    scale_y_continuous(labels = scales::comma) +
    theme(plot.title = element_text(size=10, face="bold"), 
          axis.title.y=element_blank(),
          axis.title.x=element_text(size=8),
          legend.title=element_blank(),
          legend.text=element_text(size=8),
          legend.position = "right") +
    scale_fill_brewer(palette="Spectral") 


mf <- mod_smart %>% 
  filter(Interval=="Minute") %>%
  mutate(Minutes=minutes(Time)) %>%
  group_by(Minutes, SmartFeatures) %>%
  summarise(Frequency= n()) %>%
    ggplot(aes(x=Minutes, y= Frequency, fill=SmartFeatures, group=SmartFeatures)) +
    geom_area(alpha=0.5) +
    geom_line(aes(ymax= Frequency, colour=SmartFeatures), position="stack", show.legend = FALSE) +
    ggtitle('Frequency of Minute-Tracked Smart Features') +
    scale_y_continuous(labels = scales::comma) +
    scale_fill_brewer(palette="Set2") +
    scale_color_brewer(palette="Set2") +
    theme(plot.title = element_text(size=10, face="bold"), 
          axis.title.y=element_blank(),
          axis.title.x=element_text(size=8),
          legend.title=element_blank(),
          legend.text=element_text(size=8),
          legend.position = "right")

plot_grid(hf, mf, nrow=2, align = "v")
```

A consistent frequency under 1,000 uses every hour was recorded for the hourly smart features. The four features that were recorded hourly saw an equal amount of use from hours '0-23'. The Sleep feature had the lowest frequency of use compared to the other four features recorded every minute. The minute frequencies of these features had over 20,000 uses steadily from minutes '0-59'. 


### Top Recommendations

**Drop Weight Features**

Bellabeat should consider removing or not developing any weight-related features since the five least used features were all related to weight. By shifting resources to other features that are actually popular among users, the company would be able to better market their products and attract potential customers. 


**Develop Features Related to Most Popular Features**

Bellabeat should create features related to the top five most used features (Heart Rate, Calories, Intensity, Steps and METs). For example, the company could develop cardio-related features such as a blood pressure recorder or pulse monitor to reflect the high use in the Heart Rate feature. These new features could then be added to existing Bellabeat products to the benefit of existing and potential customers. The Time wellness watch would be the perfect product to track these new features.


**Track Features by Minutes/Seconds**

The company should ensure that their features can be tracked by the minutes or seconds as the highest use were recorded during these time intervals. This was illustrated when analyzing the highly consistent minute frequency of smart features and the Minute-Wide features. If Bellabeat develops products to track features at these time intervals, it should match the high use seen during the month while also appealing to potential customers. 
